% ============================================================================
%   Claim safe AAAI Draft for the DST-EDL Architecture
%   Uncertainty to Topology: Adaptation for Structured Sparse Evidential Learning under Extreme Class Imbalance
%
%   File: dst_edl_report_en.tex
%   Draft label: tradeoff aware reported configuration (06/2026)
%   Language: English
% ============================================================================

\documentclass[letterpaper]{article}
\usepackage[submission]{aaai2026}
\usepackage{times}
% AAAI-26 forbids fontenc; keep the default font encoding required by aaai.sty.
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsfonts,bm}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{xcolor}
\usepackage{caption}
\usepackage[round]{natbib}
% AAAI-26 forbids hyperref; citations remain author--year through natbib below.
\usepackage{helvet}
\usepackage{courier}
% \setcitestyle{authoryear,round,semicolon,aysep={}}
\AtBeginDocument{\let\cite\citep}
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS

\setcounter{secnumdepth}{2}
\usepackage{xcolor}
\usepackage{cuted}
\usepackage{capt-of}

\captionsetup{font=small,labelfont=bf}
\setlength{\parindent}{1em}
\setlength{\bibsep}{2pt plus 1pt}

% ---- lightweight math shortcuts ----
\newcommand{\sg}{\mathrm{sg}}                % stop-gradient
\newcommand{\ue}{u_{e}}
\newcommand{\ua}{u_{a}}
\newcommand{\Salpha}{S_{\alpha}}
\newcommand{\maskM}{\mathbf{M}}
\newtheorem{proposition}{Proposition}
\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Uncertainty-Guided 2:4 Dynamic Sparse Evidential Learning for Rare-Event Screening)
/Author (Anonymous Authors)
}
%%%%%%%%%%
 % DO NOT CHANGE THIS

% Packages required by the manuscript body, kept separate from the protected
% AAAI configuration above to avoid reloading packages with conflicting options.
\usepackage[english]{babel}
\usepackage{adjustbox}
\usepackage{placeins}
\usepackage[final]{microtype}

\newcolumntype{C}{>{\centering\arraybackslash}c}

% --- Custom commands ---
\newcommand{\Dir}{\text{Dir}}
\newcommand{\KL}{\text{KL}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Norm}{\text{Norm}}
\newcommand{\softplus}{\text{Softplus}}
\newcommand{\pospart}[1]{\left[#1\right]_+}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\TopTwo}{Top2}

\tolerance=1500
\hyphenpenalty=50
\exhyphenpenalty=50
\lefthyphenmin=2
\righthyphenmin=3
\emergencystretch=1em
\hyphenation{ev-i-den-tial Di-rich-let un-cer-tain-ty clas-si-fi-ca-tion cal-i-bra-tion struc-tur-al top-ol-o-gy spec-i-fic-i-ty sen-si-tiv-i-ty crys-tal-li-za-tion rep-re-sen-ta-tion-al im-bal-ance com-pat-i-bil-i-ty ac-cel-er-a-tion de-ploy-ment fea-si-bil-i-ty mor-phol-o-gy con-cen-tra-tion reg-u-lar-i-za-tion}

\newtheorem{remark}{Remark}

% Combined manuscript build: retain the appendix in this source and PDF.
% Switch to \withappendixfalse only when preparing a page-limited main-paper PDF.
\newif\ifwithappendix
\withappendixtrue


% ============================================================================
\title{Uncertainty-Guided 2:4 Dynamic Sparse Evidential Learning for Rare-Event Screening}
\author{Anonymous Authors}

\begin{document}

\maketitle

\begin{abstract}
Rare-event screening requires ranking scarce positive cases in a high-sensitivity regime while representing weak or conflicting evidence under a deployable sparsity constraint. Existing imbalance-aware evidential methods modify losses, priors, margins, or decision rules but retain fixed connectivity, whereas dynamic sparse training adapts connectivity using magnitude or task-gradient signals that are agnostic to the model's evidential state. We introduce \textbf{DST-EDL}, a two-loop framework that uses Dirichlet vacuity and expected categorical entropy for active-link pruning and KL-to-uniform influence for dormant-link regrowth under exact 2:4 sparsity. A virtual effective-weight probe exposes structural gradients at masked coordinates without applying ordinary task updates to dormant parameters. On one patient-grouped ISIC 2024 split with three training seeds, DST-EDL obtains the highest observed mean $\mathrm{pAUC}_{0.80}=0.1172$ and Macro AUROC $=0.8853$ among the compared systems while preserving exact 2:4 validity, with tradeoffs in balanced accuracy and calibration. Under the base evidential objective, its topology controller exceeds Static 2:4 EDL but not RigL-style 2:4; within full-system ablations, removing dormant-link regrowth produces the largest degradation. These results support evidential state as a feasible topology-control signal in the studied screening regime rather than a universal long-tailed solution.
\end{abstract}

\section{Introduction}\label{sec:intro}

Rare-event screening is not adequately characterized by conventional imbalanced-classification accuracy. A useful screening model must rank very few positive cases in the high-sensitivity region, indicate when its evidence is weak or conflicting, and satisfy deployment constraints without obscuring signals associated with rare cases. ISIC 2024 makes this tension concrete: malignant images occur at roughly one case per thousand, while the primary metric emphasizes the part of the ROC curve above $0.80$ sensitivity \cite{isic2024slice3d}. Under this prevalence skew, majority predictions can dominate default-threshold accuracy, making ranking, calibration, and operating-point selection central to evaluation.

Existing methods address different parts of this problem. Long-tailed learning changes sampling, class weights, margins, logits, classifier heads, or representations to reduce majority-class dominance \cite{cui2019class,cao2019learning,menon2021long,ren2020balanced,kang2019decoupling}. EDL instead maps network outputs to Dirichlet evidence, yielding predictive probabilities and evidence-derived uncertainty proxies in one forward pass \cite{sensoy2018evidential}. Class-Balanced EDL uses class-level pooling and a distribution-aware learnable prior for medical imbalance \cite{xia2024classbalanced}; reweighted-margin EDL changes class emphasis and the decision margin for cervical OCT \cite{zhu2025reweighted}; and Flexible EDL expands the evidential family used to represent uncertainty \cite{yoon2025flexible}. These methods adapt the objective, prior, margin, or uncertainty representation, but retain fixed learned parameter connectivity.

Recent work also treats uncertainty as an active training-time signal. UGPL directs detailed processing toward ambiguous CT regions \cite{venkatraman2025ugpl}; TMLC uses evidential opinions for multi-view aggregation and uncertainty-guided pseudo-data generation in long-tailed learning \cite{tang2026tmlc}; and D-EDL revises evidential restriction to avoid over-penalizing ambiguous in-distribution medical samples during OOD detection \cite{fu2026dedl}. Thus uncertainty can guide data generation, spatial computation, regularization, or decisions. These closest approaches still do not use evidential state to score individual parameter connections.

Dynamic sparse training (DST) provides the complementary mechanism of changing active connections under a fixed budget. SET uses magnitude pruning and random growth, SNFS uses momentum-like signals, RigL uses supervised-loss gradients for regrowth, SRigL imposes structured fan-in, and recent permutation-augmented structured DST improves the expressivity of hardware-oriented patterns \cite{mocanu2018scalable,dettmers2019sparse,evci2020rigging,lasby2023srigl,tyagi2025padst}. These methods derive topology from parameter values, supervised optimization state, momentum, or predetermined structural layouts; they do not condition link scoring on the model's evidential state. This leaves a missing interface: can evidence-derived state guide active and dormant connections while every topology refresh remains feasible under a deployable sparsity constraint?

We introduce DST-EDL to study this interface under exact 2:4 sparsity. A mini-batch prediction loop updates active weights under a cached mask. A slower structure loop uses an ambiguity-over-vacuity derivative to score active-link removal and KL-to-uniform influence to score dormant-link regrowth. Its virtual effective-weight probe exposes structural gradients at zero-valued effective coordinates without applying ordinary task updates to dormant parameters. A bounded latent state integrates both signals before blockwise top-two projection. In line with recent critiques and density-informed reformulations of EDL \cite{shen2024mirage,carlotti2026density}, these quantities are operational structural proxies, not calibrated Bayesian posterior uncertainties.

On one patient-grouped ISIC split and three model seeds, the complete system achieves the largest observed mean $\mathrm{pAUC}_{0.80}$ and Macro AUROC among the evaluated softmax, dense evidential, and sparse evidential systems. The gain is specific to high-sensitivity ranking: several baselines retain better balanced accuracy, AP, or calibration. Under the base evidential objective, the topology controller exceeds Static 2:4 EDL but not RigL-style 2:4, so the principal table is an end-to-end system comparison rather than a causal estimate of topology alone. A separate TensorRT experiment establishes sparse-path execution and same-graph acceleration; it does not substitute deployment checkpoints for the predictive comparison.

\paragraph{Main Contributions.}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item \textbf{Evidence-conditioned topology control.} We formulate signed ambiguity-over-vacuity pruning and KL-to-uniform dormant-link influence, followed by exact blockwise 2:4 projection.
    \item \textbf{Decoupled dormant-link scoring.} A virtual effective-weight probe computes structural-objective gradients through masked coordinates without applying supervised parameter updates to dormant weights.
    \item \textbf{Bounded empirical characterization.} Three-seed ISIC comparisons, component diagnostics, an external-domain OOD diagnostic, and a separate TensorRT sparse-backend test identify both the ranking benefit and its accuracy, calibration, and deployment boundaries.
\end{itemize}

% The overview figure is intentionally omitted from the submission build.
\iffalse
\begin{figure*}[t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  font=\footnotesize,
  box/.style={draw, rounded corners=2pt, align=center, minimum height=10mm,
              text width=2.15cm, inner xsep=3pt, inner ysep=3pt},
  wide/.style={box, text width=2.55cm},
  arr/.style={-{Latex[length=2mm]}, line width=0.7pt},
  state/.style={arr, dashed},
  structural/.style={arr, densely dashed, orange!80!black}
]
% Weight-update clock (each labelled mini-batch)
\node[anchor=west, font=\bfseries\small, text=blue!65!black] at (-0.1,2.25)
  {Weight-update clock: every mini-batch; $\bm M^t$ is fixed};
\node[box, fill=gray!10]   (wb) at (1.15,1.35) {Labeled batch\\$(\bm x_b,\bm y_b)$};
\node[wide, fill=blue!9]   (wf) at (4.05,1.35) {Masked forward\\$\bm W_{\rm eff}^{t,b}$\\$=\bm W^{t,b}\odot\bm M^t$};
\node[box, fill=green!10]  (we) at (6.95,1.35) {EDL outputs\\$\bm\alpha_b,\hat{\bm p}_b,u_{e,b},u_{a,b}$};
\node[wide, fill=blue!7]   (wl) at (9.85,1.35) {Training objective\\$L_{\rm train}(\bm\alpha_b,\bm y_b)$};
\node[wide, fill=violet!9] (wo) at (12.85,1.35) {Masked AdamW\\$\nabla_{\bm W}L=$\\$\nabla_{\bm W_{\rm eff}}L\odot\bm M^t$};
\node[box, fill=gray!10]   (ww) at (15.70,1.35) {Persistent output\\$\bm W^{t,b+1}$\\mask unchanged};
\draw[arr] (wb)--(wf); \draw[arr] (wf)--(we); \draw[arr] (we)--(wl);
\draw[arr] (wl)--(wo); \draw[arr] (wo)--(ww);

% Structural clock (one amortized refresh)
\node[anchor=west, font=\bfseries\small, text=orange!80!black] at (-0.1,-0.15)
  {Structure-update clock: once per post-warmup epoch; no optimizer step on $\bm W$};
\node[box, fill=gray!10]    (sb) at (1.15,-1.10) {Natural-frequency\\proxy batch $\mathcal B$};
\node[wide, fill=orange!10] (sp) at (4.05,-1.10) {Virtual forward; BN frozen\\$\widetilde{\bm W}_{\rm eff}\leftarrow$\\$\bm W^t\odot\bm M^t$};
\node[box, fill=green!10]   (se) at (6.95,-1.10) {Proxy EDL outputs\\$\bm\alpha,u_e,u_a$};
\node[wide, fill=orange!12] (ss) at (9.85,-1.10) {Structural objectives\\$R_{\mathcal B}\!\to C$; $L_{\rm grow}\!\to G$};
\node[wide, fill=violet!10] (sv) at (12.85,-1.10) {$\Delta\bm V$; EMA; layer mean; clamp\\output $\bm V^{t+1}$};
\node[box, fill=yellow!18]  (sm) at (15.70,-1.10) {Blockwise top-two\\output $\bm M^{t+1}$\\exact 2:4};
\draw[structural] (sb)--(sp); \draw[structural] (sp)--(se); \draw[structural] (se)--(ss);
\draw[structural] (ss)--(sv); \draw[structural] (sv)--(sm);

% State hand-off between the two clocks
\node[wide, fill=gray!14] (next) at (18.20,0.12)
  {Following epoch state\\$\bm W^{t+1},\bm M^{t+1}$\\forward uses $\bm W^{t+1}\odot\bm M^{t+1}$};
\draw[state] (ww.east) -| node[pos=.68,above]{persistent weights} (next.north);
\draw[state] (sm.east) -| node[pos=.68,below]{cached mask} (next.south);
\end{tikzpicture}}
\caption{Complete two-clock DST-EDL dataflow after dense warmup; at the first refresh, $\bm V$ is initialized from $|\bm W|$. The upper lane consumes labeled mini-batches and updates only persistent active weights under a cached mask. The lower lane uses a separate natural-frequency proxy batch and virtual effective-weight tensor; its gradients are detached after producing pruning score $C$ and regrowth score $G$. Latent-state projection changes only the mask. The resulting $\bm W^{t+1}$ and $\bm M^{t+1}$ compose the sparse model used by the following epoch.}
\label{fig:workflow}
\end{figure*}
\fi

\section{Related Work}\label{sec:related}

\paragraph{Evidential learning under imbalance and shift.} Standard EDL maps a network output to Dirichlet concentrations, giving a predictive mean and evidence-derived quantities in one forward pass \cite{sensoy2018evidential}. Posterior Networks use feature-density pseudo-counts, Fisher EDL uses Fisher information, R-EDL relaxes nonessential subjective-logic settings, and Flexible EDL generalizes the evidential distribution \cite{charpentier2020posterior,deng2023fisher,chen2023redl,yoon2025flexible}. For imbalance, Class-Balanced EDL uses pooling and a learnable prior, reweighted-margin EDL changes class emphasis and margins, while TLC and AREO respectively adapt expert fusion and robust optimization \cite{xia2024classbalanced,zhu2025reweighted,li2022tlc,sapkota2023areo}. DIP-EDL further separates class prediction from input-density pseudo-counts \cite{carlotti2026density}. These methods alter evidence parameterization, objectives, priors, margins, or decisions rather than parameter connectivity.

\paragraph{Uncertainty-guided adaptation.} Uncertainty increasingly controls training-time computation rather than merely being reported. UGPL identifies ambiguous image regions and refines them locally \cite{venkatraman2025ugpl}; TMLC combines evidential opinion aggregation with uncertainty-guided pseudo-data generation \cite{tang2026tmlc}; and D-EDL replaces conventional KL restriction with a differential mechanism for medical OOD detection \cite{fu2026dedl}. These approaches adapt spatial processing, data, or regularization. DST-EDL instead uses evidential state to score active and dormant parameter coordinates during sparse training.

\paragraph{Sparse topology adaptation.} Probabilistic compression uses priors, dropout rates, or stochastic gates, while Dirichlet pruning learns distributions over removable channels or neurons \cite{louizos2017bayesian,molchanov2017variational,louizos2018learning,adamczewski2021dirichlet}. DST changes connectivity during optimization: SET uses random growth, SNFS uses momentum-like signals, RigL uses supervised gradients, SRigL constrains fan-in, and BiDST optimizes masks through a bilevel objective \cite{mocanu2018scalable,dettmers2019sparse,evci2020rigging,lasby2023srigl,ji2024bidst}. Recent work additionally freezes parameters or learns permutations for structured patterns \cite{li2025pffdst,tyagi2025padst}. DST-EDL differs at the saliency interface: an isolated evidential objective and virtual effective-weight tensor score masked coordinates before exact 2:4 projection. This does not claim that evidential saliency universally dominates task-gradient rules.

\paragraph{2:4 Execution.} Two active values per eligible four-element group can invoke NVIDIA Sparse Tensor Core kernels only when layout, dtype, shape, compression, and operator coverage are supported \cite{mishra2021accelerating,nvidiacusparselt}. Mask validity therefore establishes structural compatibility; realized speed requires backend measurement. We report the two forms of evidence separately.


\section{Preliminaries: Dirichlet Evidential Signals}\label{sec:math}

For logits $\bm{z}\in\R^K$, DST-EDL maps each class logit to non-negative evidence using Softplus,
\begin{equation}\label{eq:softplus}
    e_c=\ln(1+\exp z_c),\qquad
    \alpha_c=e_c+1,\qquad
    S=\sum_{c=1}^K\alpha_c .
\end{equation}
Softplus is used instead of an output ReLU. ReLU can create zero gradient regions for negative evidence logits. The predictive mean is $\hat p_c=\alpha_c/S$. With uniform Subjective Logic base rate $a_c=1/K$ \cite{josang2016subjective}, this equals
\begin{equation}\label{eq:p_hat}
\begin{aligned}
    \hat p_c&=b_c+a_cu_e=\frac{e_c+1}{S},\\
    b_c&=\frac{e_c}{S},\qquad
    u_e=\frac{K}{S}.
\end{aligned}
\end{equation}
Imbalance is handled through the loss, leaving the Dirichlet prior unchanged. The identity $\alpha_c=e_c+1$ and the lower bound $S\ge K$ are also preserved.

Following recent critiques of Evidential Deep Learning~\cite{shen2024mirage}, we treat EDL signals as operational uncertainty proxies, not exact Bayesian posterior quantities. Within this framework, Dirichlet vacuity $u_e=K/S$ serves as a low evidence structural proxy and expected categorical entropy, formulated as:
\begin{equation}\label{eq:u_a}
    u_a = \sum_{c=1}^K \frac{\alpha_c}{S} \left[\psi(S + 1) - \psi(\alpha_c + 1)\right]
\end{equation}
as an ambiguity proxy. In implementation, $u_a$ is clamped below at zero to ensure numerical stability against floating point underflow.

For finite logits, Softplus ensures $\alpha_c>1$; formal validity statements are provided in the \ifwithappendix Appendix~\ref{app:proofs}\else supplementary material\fi.

\section{DST-EDL Framework}\label{sec:methodology}

The framework has two state transitions. The weight loop maps a labeled mini-batch and fixed mask to updated persistent weights $\bm W^{t+1}$. The structure loop maps an unlabeled proxy batch and the current effective model to latent scores $\bm V^{t+1}$ and a refreshed mask $\bm M^{t+1}$. Both operate on the same layerwise persistent tensor $\bm W^{(l)}$; the auxiliary probe introduced below is a temporary copy, not a second set of trainable weights.

\subsection{Prediction and Weight-Update Loop}\label{sec:weight_loop}
For mini-batch $(\bm x,\bm y)$, layer $l$ uses the effective tensor
\begin{equation}\label{eq:wrapped_weights}
    \bm{W}_{\text{eff}}^{(l)}=\bm{W}^{(l)}\odot\bm{M}^{(l)} .
\end{equation}
The EDL head returns $(\bm\alpha,\hat{\bm p},u_e,u_a)$ from Sec.~\ref{sec:math}; the labeled objective in Sec.~\ref{sec:training_setup} then updates $\bm W$ while $\bm M$ is cached. For any differentiable mini-batch loss,
\begin{equation}\label{eq:masked_gradient}
    \nabla_{\bm W^{(l)}}L_{\rm train}
    =\nabla_{\bm W_{\rm eff}^{(l)}}L_{\rm train}\odot\bm M^{(l)}.
\end{equation}
Thus ordinary gradient, weight decay, and optimizer-moment updates are restricted to active coordinates. The loop outputs updated persistent weights but does not alter $\bm V$ or $\bm M$.

\subsection{Evidential Structure-Update Loop}\label{sec:structure_loop}
Once per post-warmup epoch, the structure loop receives a natural-frequency proxy batch $\mathcal B$ and the current effective model. Unlike the weight loop, it produces saliency scores and a new mask but never applies its gradients to persistent weights. Structural scoring uses a dense auxiliary tensor initialized at the current masked values:
\begin{equation}\label{eq:virtual_probe}
    \widetilde{\bm W}_{\mathrm{eff}}\leftarrow\bm W\odot\bm M .
\end{equation}
The probe makes the auxiliary objective differentiable with respect to every effective coordinate, including a coordinate whose current value is zero because it is masked. Its gradient is detached after scoring; only later mask projection can reactivate a dormant weight. Unlike conventional DST, the topology scores arise from isolated evidential objectives rather than the supervised task loss. The same proxy batch is reused for pruning and regrowth with BatchNorm statistics frozen; construction and tie-handling details are given in Appendix~\ref{app:algorithm}.

% The DST comparison figure is intentionally omitted from the submission build.
\iffalse
\begin{figure*}[t]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  font=\footnotesize,
  box/.style={draw, rounded corners=2pt, align=center, text width=2.55cm,
              minimum height=16mm, inner xsep=3pt, inner ysep=3pt},
  rigl/.style={-{Latex[length=2mm]}, line width=0.75pt, blue!65!black},
  ours/.style={-{Latex[length=2mm]}, line width=0.75pt, orange!85!black}
]
% Objective-matched topology-update comparison
\node[anchor=west, font=\bfseries\small, text=blue!65!black] at (-0.05,2.25)
  {RigL-style 2:4: topology saliency comes from the supervised task objective};
\node[box, fill=gray!10]  (r0) at (1.35,1.20) {Inputs\\labeled batch $(\bm x,\bm y)$\\current $\bm W,\bm M$};
\node[box, fill=blue!8]  (r1) at (4.35,1.20) {Task forward/backward\\$L_{\rm task}$\\virtual task gradient $g^{\rm task}$};
\node[box, fill=blue!10] (r2) at (7.35,1.20) {Active-link removal\\magnitude score\\$|W_{ij}|$};
\node[box, fill=blue!10] (r3) at (10.35,1.20) {Dormant-link growth\\$H_{ij}=\mathrm{Rank}_l$\\$\bigl|g^{\rm task}_{ij}\bigr|$};
\node[box, fill=violet!9] (r4) at (13.35,1.20) {Blockwise candidate update\\task-based keep / replace ordering};
\node[box, fill=yellow!18] (r5) at (16.35,1.20) {Top-two projection\\$\bm M^{t+1}\in\mathcal C_{2:4}$};
\draw[rigl] (r0)--(r1); \draw[rigl] (r1)--(r2); \draw[rigl] (r2)--(r3);
\draw[rigl] (r3)--(r4); \draw[rigl] (r4)--(r5);

\node[anchor=west, font=\bfseries\small, text=orange!85!black] at (-0.05,-0.05)
  {DST-EDL: topology saliency comes from isolated evidential structural objectives};
\node[box, fill=gray!10]   (d0) at (1.35,-1.65) {Inputs\\proxy batch $\mathcal B$\\current $\bm W,\bm M,\bm V$};
\node[box, text width=2.78cm, fill=orange!10] (d1) at (4.35,-1.65) {Virtual probe; BN frozen\\$\widetilde{\bm W}_{\rm eff}\leftarrow\bm W\odot\bm M$\\no optimizer step};
\node[box, text width=2.78cm, fill=green!10]  (d2) at (7.35,-1.65) {Evidential objectives\\$R_{\mathcal B}=\mathbb E[u_a/(u_e+\epsilon)]$\\$L_{\rm grow}=\mathrm{KL}(\Dir(\bm\alpha)\Vert\Dir(\bm1))$};
\node[box, text width=2.78cm, fill=orange!12] (d3) at (10.35,-1.65) {Detached virtual gradients\\$C=\mathrm{Rank}([Wg^R]_+)$\\$G=\mathrm{Rank}(|g^K|)$};
\node[box, text width=2.78cm, fill=violet!10] (d4) at (13.35,-1.65) {Latent-state update\\active: $G-\lambda_pC$\\dormant: $\rho G+\Xi$\\EMA, mean center, clamp};
\node[box, fill=yellow!18] (d5) at (16.35,-1.65) {Top-two projection\\$\bm V^{t+1}\mapsto\bm M^{t+1}$\\$\bm M^{t+1}\in\mathcal C_{2:4}$};
\draw[ours] (d0)--(d1); \draw[ours] (d1)--(d2); \draw[ours] (d2)--(d3);
\draw[ours] (d3)--(d4); \draw[ours] (d4)--(d5);
\end{tikzpicture}}
\caption{Like-for-like topology-refresh comparison under the same exact 2:4 output constraint. RigL-style 2:4 removes active links by magnitude and ranks dormant candidates with gradients of the supervised task loss. DST-EDL instead evaluates a separate natural-frequency proxy batch and uses detached virtual-coordinate gradients $g^R=\nabla_{\widetilde W}R_{\mathcal B}$ and $g^K=\nabla_{\widetilde W}L_{\rm grow}$ to construct $C$ and $G$. A bounded latent state integrates these scores before top-two projection. In both rows, ordinary supervised optimization of persistent weights remains separate from topology refresh.}
\label{fig:dst_comparison}
\end{figure*}
\fi

For pruning, the proxy batch risk ratio is
\begin{equation}\label{eq:batch_risk}
    R_{\mathcal{B}}
    =
    \frac{1}{|\mathcal{B}|}
    \sum_{n\in\mathcal{B}}
    \frac{u_a^{(n)}}{u_e^{(n)}+\epsilon}.
\end{equation}
The ratio increases with ambiguity and attenuates removal pressure when evidence is scarce. Because it can decrease through either reduced ambiguity or increased vacuity, it is interpreted only as a local removal heuristic, not a descent objective. Active-link saliency is
\begin{equation}\label{eq:C_ij}
\begin{aligned}
    g^R_{ij}
    &:=\frac{\partial R_{\mathcal B}}
    {\partial\widetilde W_{\mathrm{eff},ij}},\\
    C_{ij}&=\operatorname{Rank}_{l}
    \!\left([W_{ij}g^R_{ij}]_+\right),
    \qquad [x]_+=\max(x,0).
\end{aligned}
\end{equation}
Here $\operatorname{Rank}_{l}$ normalizes positive saliencies within layer $l$; the positive part retains only removals locally predicted to reduce the ratio. This is a first-order ranking rule, not an optimality guarantee.

For regrowth, the reported setting uses KL uniform regrowth:
\begin{equation}\label{eq:grow_objective}
    L_{\text{grow}}
    =
    \frac{1}{|\mathcal{B}|}\sum_{n\in\mathcal{B}}
    \KL\!\left(
    \Dir(\bm{\alpha}^{(n)})\middle\|\Dir(\mathbf{1})
    \right).
\end{equation}
Its absolute probe gradient rank defines dormant link influence:
\begin{equation}\label{eq:G_ij}
\begin{aligned}
    g^K_{ij}
    &:=\frac{\partial L_{\mathrm{grow}}}
    {\partial\widetilde W_{\mathrm{eff},ij}},\\
    G_{ij}&=\operatorname{Rank}_{l}\!\left(|g^K_{ij}|\right).
\end{aligned}
\end{equation}
This ranks coordinates by local influence on departure from the fixed uninformative reference $\Dir(\mathbf1)$. We retain that reference so saliency remains comparable across topology refreshes; imbalance is handled in the supervised objective. Recent analysis of KL regularization in medical EDL cautions that it can over-penalize ambiguous in-distribution samples when used as an uncertainty restriction \cite{fu2026dedl}; here the absolute gradient is therefore an influence magnitude, not a direction, affected class, exact epistemic quantity, or guaranteed discrimination benefit.

The latent score update separates active retention from dormant reactivation. First, we compute the raw structural step $\Delta V_{ij}^{(t)}$:
\begin{equation}\label{eq:score_step}
\begin{aligned}
    A_{ij}^{(t)} &= G_{ij}^{(t)}-\lambda_p C_{ij}^{(t)},\\
    D_{ij}^{(t)} &= \rho G_{ij}^{(t)}+\Xi_{ij}^{(l,t)},\\
    \Delta V_{ij}^{(t)}
    &=M_{ij}^{(t)}A_{ij}^{(t)}\\
    &\quad+\left(1-M_{ij}^{(t)}\right)D_{ij}^{(t)}.
\end{aligned}
\end{equation}
Active coordinates combine regrowth influence with signed removal pressure; dormant coordinates receive attenuated regrowth influence. The nonnegative tie-breaker $\bm\Xi^{(l,t)}$ is zero unless the entire layer has numerically flat regrowth saliency, $\max_{ij}G_{ij}^{(l,t)}\le\tau$; its exact vacuity-rank modulation is given in Appendix~\ref{app:algorithm}. An exponential moving average then smooths the update, after which layerwise mean centering prevents global drift and clamping bounds the state:
\begin{equation}\label{eq:score_update}
\begin{aligned}
    \mu_{ij}^{(t+1)} &= \beta_m\mu_{ij}^{(t)} +(1-\beta_m)\Delta V_{ij}^{(t)},\\
    X_{ij}^{(t+1)} &= V_{ij}^{(t)}+\eta_{\text{struct}}\mu_{ij}^{(t+1)},\\
    \bar X_l^{(t+1)}
    &=\operatorname{mean}_{l}(\bm X^{(t+1)}),\\
    V_{ij}^{(t+1)}
    &=\operatorname{clamp}\!\bigl(
    X_{ij}^{(t+1)}-\bar X_l^{(t+1)},\\
    &\quad -V_{\max},V_{\max}\bigr).
\end{aligned}
\end{equation}
The structure loop therefore outputs only $\bm V^{t+1}$. Each refresh adds one proxy forward pass and two scalar backward passes, amortized over the following epoch. No wall-clock training-overhead reduction is claimed.

\subsection{Coupling the Loops in a Sparse Backbone}\label{sec:loop_coupling}
The structure output becomes a valid mask by projecting scores independently within every eligible four-element block:
\begin{equation}\label{eq:c24}
\begin{aligned}
\mathcal C_{2:4}
&=\left\{\bm M\in\{0,1\}^{d_{\rm out}\times d_{\rm in}}:\right.\\
&\left.\quad\sum_{j=0}^{3}M_{i,4k+j}=2\ \forall i,k\right\}.
\end{aligned}
\end{equation}
Writing $I_{i,k}^{(l)}=\TopTwo_{j\in\{0,1,2,3\}}V_{i,4k+j}^{(l)}$, the refresh is
\begin{equation}\label{eq:mask}
\begin{aligned}
M_{i,4k+j}^{(l),t+1}
&=\mathbb 1[j\in I_{i,k}^{(l)}],\\
\bm M^{t+1}
&=\Pi_{\mathcal C_{2:4}}(\bm V^{t+1}).
\end{aligned}
\end{equation}
The weight loop supplies $\bm W^{t+1}$ and the structure loop supplies $\bm M^{t+1}$; their only coupling is the next effective tensor $\bm W_{\rm eff}^{t+1}=\bm W^{t+1}\odot\bm M^{t+1}$. This identity links the persistent weights in Eq.~\ref{eq:wrapped_weights} to every pruning, regrowth, and projection step: $\widetilde{\bm W}_{\rm eff}$ is discarded after scoring, whereas $\bm W$ persists across mask changes. Reactivated coordinates resume from their stored value.

We wrap eligible $3\times3$ residual convolutions, retain the dense RGB stem and classifier, and leave incomplete four-element groups dense. Warmup and layout details are given in Appendix~\ref{app:algorithm}. Thus labeled data update $\bm W$, proxy data update $\bm V$ and $\bm M$, and their composition supplies the next sparse EDL forward pass.

\section{Training Setup}\label{sec:training_setup}

Active weights are trained with the expected Dirichlet classification loss \cite{sensoy2018evidential}, square-root class weighting, and stop-gradient focal modulation \cite{lin2017focal}. This imbalance-aware objective is part of the evaluated system rather than a separate loss contribution:
\begin{equation}\label{eq:efl}
\begin{aligned}
    L_{\rm train}\equiv L_{\rm EFL}
    &=\frac{1}{N}\sum_{n=1}^N\mathcal F_n(t)\\
    &\quad\times\left[
    \bar\omega_{y_n}\ell_n
    +\lambda_{\mathrm{KL}}\mu(t)k_n\right],\\
    \ell_n&=\psi(S_n)-\psi(\alpha_{n,y_n}),\\
    k_n&=\KL\!\left(
    \Dir(\widetilde{\bm\alpha}_n)\parallel\Dir(\mathbf1)\right).
\end{aligned}
\end{equation}
Its components are
\begin{equation}\label{eq:sample_weight}
\bar\omega_c=\min\!\left(
\sqrt{N_{\rm major}/N_c},\omega_{\max}\right),
\end{equation}
and the remaining parameters are defined by:
\begin{equation}\label{eq:kl_loss_main}
\begin{aligned}
\mathcal F_n(t)&=\left(1-\sg[\hat p_{n,y_n}]\right)^{\gamma(t)},\\
\widetilde\alpha_{n,c}&=y_{n,c}+(1-y_{n,c})\alpha_{n,c},\\
\mu(t)&=\min(1,t/t_{\rm anneal}).
\end{aligned}
\end{equation}
Square-root weighting is less extreme than inverse-frequency weighting, and the cap bounds minority amplification. Stop gradient lets focal modulation change sample emphasis without creating another derivative path through confidence. The reported row uses symmetric KL shrinkage; annealing, focal scheduling, and the asymmetric diagnostic are specified in Appendix~\ref{app:algorithm}. After Eq.~\ref{eq:softplus}, evidence is clipped at 20 in implementation for numerical stability. AdamW uses a one-epoch learning-rate warmup followed by cosine decay over 40 epochs; evidential losses are evaluated in FP32 under mixed-precision training.

\section{Experimental Setup}\label{sec:experiment}

\subsection{Data and Evaluation Protocol}
ISIC 2024 contains 3D total-body-photography lesion crops with below $0.12\%$ malignant prevalence \cite{isic2024slice3d}. A split seed of 42 fixes one patient-grouped 70/5/5/20 train/validation/calibration/test partition. Only the training partition is majority-subsampled to 1:20, giving 5,775 training images; validation (21,930), calibration (24,235), and test (65,931) retain natural prevalence. The test set contains 77 malignant and 65,854 benign images. Validation selects the decision threshold, calibration fits bias and temperature, and test data are used only after configuration choices are frozen. Images are resized to $224\times224$ with ImageNet normalization.

The \emph{split seed} determines patient membership once. The three \emph{training seeds} (42, 123, 456) keep that split fixed but vary initialization, mini-batch order, augmentation, and stochastic topology operations. Reported standard deviations therefore measure training-seed variation, not alternative cohorts or splits. Each configuration uses the final-epoch checkpoint under the same 40-epoch schedule. The validation operating point is
\begin{equation}\label{eq:threshold}
\tau^*=\argmax_{\tau}\;\mathrm{TNR}_{\rm val}(\tau)
\quad\mathrm{s.t.}\quad \mathrm{TPR}_{\rm val}(\tau)\ge0.80,
\end{equation}
and is frozen before test evaluation.

\paragraph{Metrics.} Let $q=1-\mathrm{TPR}$, let $B_b$ be one of 15 adaptive equal-frequency calibration bins, and let $(P_r,R_r)$ denote precision and recall along ranked predictions. We use
\begin{equation}\label{eq:metrics}
\begin{aligned}
\mathrm{BalAcc}&=\tfrac12(\mathrm{TPR}+\mathrm{TNR}),\\
\mathrm{pAUC}_{0.80}&=\int_{0}^{0.20}\mathrm{TNR}(q)\,dq,\\
\mathrm{M\!\text{-}AUROC}&=\tfrac1K\sum_{c=1}^{K}\mathrm{AUROC}_c,\\
\mathrm{AP}&=\sum_r(R_r-R_{r-1})P_r,\\
\mathrm{ECE}&=\sum_b\tfrac{|B_b|}{N}
\left|\mathrm{acc}(B_b)-\mathrm{conf}(B_b)\right|.
\end{aligned}
\end{equation}
BalAcc gives equal weight to both classes; unnormalized pAUC$_{0.80}\in[0,0.20]$ measures specificity in the required high-sensitivity region; Macro AUROC measures threshold-independent discrimination; average precision (AP, computed by \texttt{average\_precision\_score}) emphasizes rare-positive retrieval; and ECE measures confidence--accuracy mismatch, including overconfidence. AURC, reported in Appendix~\ref{app:seed_level}, integrates selective-prediction risk over coverage (lower is better).

\subsection{Baselines, Backbone, and Scope}
The shared ResNet-18 protocol includes ordinary softmax baselines (CE, LDAM-DRW, and MiSLAS-style LAS+cRT), dense evidential baselines (standard, Fisher, Flexible, and R-EDL), Static 2:4 EDL, RigL-style 2:4, and DST-EDL. All predictive rows share the patient split, training subsample, ImageNet initialization, architecture, epoch budget, checkpoint rule, calibration family, and threshold selection. They do \emph{not} share one objective: the dense/Static/RigL evidential rows use the base symmetric EDL loss, whereas DST-EDL uses Eq.~\ref{eq:efl}. Table~\ref{tab:main_evidential_comparison} is therefore an end-to-end comparison. The base-loss controller row and one-component removals diagnose attribution but do not estimate a topology$\times$focal interaction.

The completed external diagnostic applies fixed ISIC-trained models to held-out PAD-UFES-20 smartphone images \cite{pacheco2020padufes}; PAD data are never used for training, calibration, or threshold selection. Its prespecified layer-3 KNN score measures source-to-external separation, not external diagnostic accuracy or an advantage of EDL uncertainty. Hardware profiling likewise uses a separately exported, layout-compatible graph; those predictive and deployment metrics are not interchanged.

\section{Results on the ISIC 2024 Case Study}\label{sec:results}

\subsection{High-Sensitivity Ranking and Tradeoffs}
Table~\ref{tab:main_evidential_comparison} broadens the comparison beyond EDL. MiSLAS-style LAS+cRT is the strongest softmax row at pAUC$_{0.80}=0.1032$, while RigL-style 2:4 is the strongest sparse evidential baseline at $0.0982$. DST-EDL obtains the largest observed means, pAUC$_{0.80}=0.1172$ and Macro AUROC $=0.8853$. The former is an area with maximum $0.20$, not an accuracy of $11.72\%$.

The ordering is not uniform. Dense EDL has higher validation-selected balanced accuracy ($0.8221$ vs.~$0.7777$), R-EDL has higher AP ($0.0422$ vs.~$0.0335$), and both dense and softmax baselines have lower ECE. The small softmax ECE values are also aided by the negative-dominated test prevalence. Thus the empirical advantage is specific to high-sensitivity and global ranking. Because all seeds reuse one split and no valid paired patient bootstrap is available, the differences remain descriptive rather than statistically established.

\begin{table}[tbp]
\centering
\caption{ISIC 2024 ranking and calibration comparison on one fixed patient split. Values are mean $\pm$ training-seed SD for seeds 42/123/456. These metrics are threshold-independent; bold marks the largest observed mean, not significance. Lower ECE is better.}
\label{tab:main_evidential_comparison}
\scriptsize
\renewcommand{\arraystretch}{1.03}
\setlength{\tabcolsep}{1.4pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l C C C C@{}}
\toprule
\textbf{Model} & \textbf{pAUC$_{0.80}$} $\uparrow$ & \textbf{M-AUROC} $\uparrow$ & \textbf{AP} $\uparrow$ & \textbf{ECE} $\downarrow$ \\
\midrule
\multicolumn{5}{l}{\textbf{Softmax long-tailed baselines}} \\
Standard CE & $0.0907\pm0.0078$ & $0.8560\pm0.0106$ & $0.0274\pm0.0018$ & $\mathbf{0.0007\pm0.0001}$ \\
LDAM-DRW & $0.0985\pm0.0063$ & $0.8623\pm0.0046$ & $0.0319\pm0.0052$ & $0.0009\pm0.0001$ \\
MiSLAS-style & $0.1032\pm0.0064$ & $0.8708\pm0.0090$ & $0.0313\pm0.0051$ & $0.0015\pm0.0001$ \\
\midrule
\multicolumn{5}{l}{\textbf{Dense evidential baselines}} \\
Dense EDL & $0.0885\pm0.0124$ & $0.8663\pm0.0102$ & $0.0413\pm0.0085$ & $0.0524\pm0.0013$ \\
Fisher EDL & $0.0842\pm0.0111$ & $0.8604\pm0.0163$ & $0.0414\pm0.0104$ & $0.0519\pm0.0009$ \\
Flexible EDL & $0.0847\pm0.0022$ & $0.8595\pm0.0078$ & $0.0373\pm0.0064$ & $0.0525\pm0.0013$ \\
R-EDL & $0.0743\pm0.0175$ & $0.8479\pm0.0241$ & $\mathbf{0.0422\pm0.0051}$ & $0.0523\pm0.0012$ \\
\midrule
\multicolumn{5}{l}{\textbf{Sparse 2:4 evidential systems (50\% wrapped density)}} \\
Static 2:4 EDL & $0.0803\pm0.0036$ & $0.8449\pm0.0051$ & $0.0377\pm0.0039$ & $0.0529\pm0.0002$ \\
RigL-style 2:4 & $0.0982\pm0.0064$ & $0.8630\pm0.0079$ & $0.0371\pm0.0089$ & $0.0550\pm0.0004$ \\
\textbf{DST-EDL} & $\mathbf{0.1172\pm0.0067}$ & $\mathbf{0.8853\pm0.0089}$ & $0.0335\pm0.0037$ & $0.0554\pm0.0007$ \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\subsection{Objective and Topology Attribution}
Under the base evidential objective, the controller without focal modulation reaches pAUC$_{0.80}=0.0898$: above Static 2:4 EDL ($0.0803$) but below RigL-style 2:4 ($0.0982$). The main comparison therefore does not isolate evidential saliency from the objective. Table~\ref{tab:controlled} provides the narrower one-component evidence. Removing regrowth produces the largest observed drop ($-0.0528$), followed by focal modulation ($-0.0270$) and the signed pruner ($-0.0117$). Removing the tie-breaking heuristic slightly raises mean pAUC from $0.1168$ to $0.1173$ in the full supplementary table, so it is not claimed as a ranking contributor. These are within-suite effects, not a factorial interaction estimate or proof that evidential regrowth dominates task-gradient alternatives.

\begin{table}[tbp]
\centering
\caption{Means from a separate three-seed component suite; SDs and remaining variants are in the supplement. Its full reference is not the Table~\ref{tab:main_evidential_comparison} checkpoint; $\Delta$ is relative to $0.1168$.}
\label{tab:controlled}
\scriptsize
\renewcommand{\arraystretch}{1.03}
\setlength{\tabcolsep}{1.4pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l C C C C@{}}
\toprule
\textbf{Variant} & \textbf{pAUC$_{0.80}$} & \textbf{M-AUROC} & \textbf{ECE} & \textbf{$\Delta$pAUC} \\
\midrule
Full suite reference & $0.1168$ & $0.8891$ & $0.0564$ & --- \\
w/o signed pruner & $0.1051$ & $0.8720$ & $0.0552$ & $-0.0117$ \\
w/o regrower & $0.0640$ & $0.7663$ & $0.0556$ & $-0.0528$ \\
w/o focal modulation & $0.0898$ & $0.8546$ & $0.0544$ & $-0.0270$ \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\subsection{Structure, Efficiency, and External Shift}
DST-EDL activates 5,598,786 of 11,177,538 effective-weight coordinates (50.09\% overall; exactly 50\% in wrapped layers), and all 2,789,376 eligible blocks satisfy 2:4. Theoretical cost falls from 1.8136 to 0.9658 GMAC/image ($-46.75\%$). On one frozen RTX A2000 TensorRT 11.1 graph at batch 64, enabling sparse tactics changes 4847.1 to 5546.3 images/s and 13.20 to 11.54 ms, a $1.14\times$ throughput gain and 12.6\% latency reduction. This backend result is separate from Table~\ref{tab:main_evidential_comparison}; it establishes sparse-path execution, not an end-to-end accuracy--latency frontier.

Fixed ISIC models separate held-out PAD-UFES-20 images by a prespecified layer-3 KNN score (Appendix~\ref{app:external}). This source-only diagnostic shows feature-space domain separation, not external lesion classification or comparative EDL-OOD superiority.

\section{Limitations and Broader Impact}\label{sec:limitations}
\paragraph{Limitations.} The empirical claim is conditional on one patient-grouped split with 77 malignant test images; three training seeds do not quantify cohort or split uncertainty. The principal comparison is end-to-end because objective-matched Static/RigL+focal rows and a Class-Balanced EDL baseline are unavailable, leaving the topology--objective interaction unidentified. DST-EDL improves ranking but not balanced accuracy, AP, or calibration uniformly, and the component suite does not compare all regrowth alternatives or record epochwise mask turnover. The PAD result uses a feature-space KNN score rather than an evidential OOD comparison. TensorRT acceleration is confined to one RTX A2000 batch-64 configuration; batch 1 shows no gain, and predictive parity of the exported graph is unverified. A negative CIFAR-100-LT diagnostic further rules out a domain-general long-tailed claim.

\paragraph{Broader Impact.} DST-EDL is evaluated as a research screening and triage model, not an autonomous diagnostic system. Clinical use would require multi-center external validation, subgroup analysis, human review of referred or deferred cases, and appropriate patient-data governance.

\section{Conclusion}\label{sec:conclusion}
We introduced DST-EDL, an uncertainty-to-topology framework combining evidence-derived structural saliency, virtual dormant-link probing, and exact 2:4 projection. On the studied ISIC protocol, the complete system improves high-sensitivity ranking over the compared systems while controlled diagnostics expose calibration tradeoffs and incomplete isolation of the topology effect. The results establish the feasibility of using evidential state for dynamic structured sparsity in this screening regime; objective-matched topology comparisons, cohort-level uncertainty, and broader long-tailed validity remain open.
\begin{thebibliography}{10}

\bibitem[Sensoy et~al.(2018)Sensoy et~al.]{sensoy2018evidential}
M.~Sensoy, L.~Kaplan, and M.~Kandemir.
\newblock Evidential deep learning to quantify classification uncertainty.
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)}, 2018.

\bibitem[Xia et~al.(2024)Xia et~al.]{xia2024classbalanced}
T.~Xia, T.~Dang, J.~Han, L.~Qendro, and C.~Mascolo.
\newblock Uncertainty-aware health diagnostics via class-balanced evidential deep learning.
\newblock {\em IEEE Journal of Biomedical and Health Informatics}, 28(11):6417--6428, 2024.

\bibitem[Pacheco et~al.(2020)Pacheco et~al.]{pacheco2020padufes}
A.~G.~C. Pacheco, G.~R. Lima, A.~S. Salom\~ao, B.~A. Krohling, I.~P. Biral,
G.~G. de~Angelo, F.~C.~R. Alves~Jr., J.~G.~M. Esgario, A.~C. Simora,
P.~B.~C. Castro, F.~B. Rodrigues, P.~H.~L. Frasson, R.~A. Krohling,
H.~Knidel, M.~C.~S. Santos, R.~B. do~Esp\'irito~Santo,
T.~L.~S.~G. Macedo, T.~R.~P. Canuto, and L.~F.~S. de~Barros.
\newblock PAD-UFES-20: A skin lesion dataset composed of patient data and clinical images collected from smartphones.
\newblock {\em Data in Brief}, 32:106221, 2020.

\bibitem[Malinin and Gales(2018)Malinin and Gales]{malinin2018predictive}
A.~Malinin and M.~Gales.
\newblock Predictive uncertainty estimation via prior networks.
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)}, 2018.

\bibitem[Evci et~al.(2020)Evci et~al.]{evci2020rigging}
U.~Evci, T.~Gale, J.~Menick, P.~S. Castro, and E.~Elsen.
\newblock Rigging the lottery: Making all tickets winners.
\newblock In {\em International Conference on Machine Learning (ICML)}, 2020.

\bibitem[Cui et~al.(2019)Cui et~al.]{cui2019class}
Y.~Cui, M.~Jia, T.-Y. Lin, Y.~Song, and S.~Belongie.
\newblock Class balanced loss based on effective number of samples.
\newblock In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}, 2019.

\bibitem[Cao et~al.(2019)Cao et~al.]{cao2019learning}
K.~Cao, C.~Wei, A.~Gaidon, N.~Arechiga, and T.~Ma.
\newblock Learning imbalanced datasets with label distribution aware margin loss.
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)}, 2019.

\bibitem[Menon et~al.(2021)Menon et~al.]{menon2021long}
A.~K. Menon, S.~Jayasumana, A.~S. Rawat, H.~Jain, A.~Veit, and S.~Kumar.
\newblock Long tail learning via logit adjustment.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2021.

\bibitem[Ren et~al.(2020)Ren et~al.]{ren2020balanced}
J.~Ren, C.~Yu, S.~Sheng, X.~Ma, H.~Zhao, S.~Yi, and H.~Li.
\newblock Balanced meta softmax for long tailed visual recognition.
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)}, 2020.

\bibitem[Kang et~al.(2020)Kang et~al.]{kang2019decoupling}
B.~Kang, S.~Xie, M.~Rohrbach, Z.~Yan, A.~Gordo, J.~Feng, and Y.~Kalantidis.
\newblock Decoupling representation and classifier for long tailed recognition.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2020.

\bibitem[Zhong et~al.(2021)Zhong et~al.]{zhong2021improving}
Z.~Zhong, J.~Cui, S.~Liu, and J.~Jia.
\newblock Improving calibration for long tailed recognition.
\newblock In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}, 2021.

\bibitem[Zhou et~al.(2020)Zhou et~al.]{zhou2020bbn}
B.~Zhou, Q.~Cui, X.-S. Wei, and Z.-M. Chen.
\newblock BBN: Bilateral branch network with cumulative learning for long tailed visual recognition.
\newblock In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}, 2020.

\bibitem[Cui et~al.(2021)Cui et~al.]{cui2021parametric}
J.~Cui, Z.~Zhong, S.~Liu, B.~Yu, and J.~Jia.
\newblock Parametric contrastive learning.
\newblock In {\em IEEE/CVF International Conference on Computer Vision (ICCV)}, 2021.

\bibitem[Deng et~al.(2023)Deng et~al.]{deng2023fisher}
D.~Deng, G.~Chen, Y.~Yu, F.~Liu, and P.-A. Heng.
\newblock Uncertainty estimation by Fisher information based evidential deep learning.
\newblock In {\em International Conference on Machine Learning (ICML)}, PMLR, 2023.

\bibitem[Chen et~al.(2024a)Chen et~al.]{chen2023redl}
M.~Chen, J.~Gao, and C.~Xu.
\newblock R-EDL: Relaxing nonessential settings of evidential deep learning.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2024.

\bibitem[Chen et~al.(2024b)Chen et~al.]{chen2024reedl}
M.~Chen, J.~Gao, and C.~Xu.
\newblock Revisiting essential and nonessential settings of evidential deep learning.
\newblock {\em arXiv preprint arXiv:2410.00393}, 2024.

\bibitem[Yoon and Kim(2025)Yoon and Kim]{yoon2025flexible}
T.~Yoon and H.~Kim.
\newblock Uncertainty estimation by flexible evidential deep learning.
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)}, 2025.

\bibitem[Zhu et~al.(2025)Zhu et~al.]{zhu2025reweighted}
H.~Zhu, Y.~Qu, Y.~Ma, and Y.~Pei.
\newblock Evidential deep learning with reweighted margin adjustment for uncertainty-driven cervical {OCT} image diagnosis.
\newblock In {\em IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP)}, pages 1--5, 2025.

\bibitem[Venkatraman et~al.(2025)Venkatraman et~al.]{venkatraman2025ugpl}
S.~Venkatraman, P.~S., R.~R. Madavan, and C.~S.
\newblock {UGPL}: Uncertainty-guided progressive learning for evidence-based classification in computed tomography.
\newblock In {\em Proceedings of the IEEE/CVF International Conference on Computer Vision Workshops (ICCVW)}, pages 969--979, 2025.

\bibitem[Fu et~al.(2026)Fu et~al.]{fu2026dedl}
W.~Fu, Y.~Chen, Y.~Liu, and X.~Yue.
\newblock {D-EDL}: Differential evidential deep learning for robust medical out-of-distribution detection.
\newblock {\em Medical Image Analysis}, 109:103888, 2026.

\bibitem[Tang et~al.(2026)Tang et~al.]{tang2026tmlc}
C.~Tang, Y.~Shi, G.~Lin, L.~Xing, and L.~Shi.
\newblock Trusted multi-view learning for long-tailed classification.
\newblock In {\em Proceedings of the AAAI Conference on Artificial Intelligence}, 40(30):25814--25822, 2026.

\bibitem[Carlotti et~al.(2026)Carlotti et~al.]{carlotti2026density}
P.~Carlotti, N.~Gligi\'c, and A.~Farahi.
\newblock Density-informed pseudo-counts for calibrated evidential deep learning.
\newblock {\em arXiv preprint arXiv:2602.01477}, 2026.

\bibitem[Yu et~al.(2024)Yu et~al.]{yu2024anedl}
Y.~Yu, D.~Deng, F.~Liu, Q.~Dou, Y.~Jin, G.~Chen, and P.-A. Heng.
\newblock ANEDL: Adaptive negative evidential deep learning for open set semi supervised learning.
\newblock In {\em AAAI Conference on Artificial Intelligence}, 38(15):16587--16595, 2024.

\bibitem[Shen et~al.(2024)Shen et~al.]{shen2024mirage}
M.~Shen, J.~J. Ryu, S.~Ghosh, Y.~Bu, P.~Sattigeri, S.~Das, and G.~W. Wornell.
\newblock Are uncertainty quantification capabilities of evidential deep learning a mirage?
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)}, 2024.

\bibitem[Dettmers and Zettlemoyer(2019)Dettmers and Zettlemoyer]{dettmers2019sparse}
T.~Dettmers and L.~Zettlemoyer.
\newblock Sparse networks from scratch: Faster training without losing performance.
\newblock In {\em ICLR Workshop}, 2019.

\bibitem[Lee et~al.(2019)Lee et~al.]{lee2019snip}
N.~Lee, T.~Ajanthan, and P.~H.~S. Torr.
\newblock SNIP: Single shot network pruning based on connection sensitivity.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2019.

\bibitem[Li et~al.(2025)Li et~al.]{li2025pffdst}
L.~Li, H.~Yang, J.~Guo, H.~Yu, M.~Qin, and T.~Zhang.
\newblock An efficient and accurate dynamic sparse training framework based on parameter freezing.
\newblock In {\em AAAI Conference on Artificial Intelligence}, 39(17):18350--18358, 2025.

\bibitem[Lasby et~al.(2024)Lasby et~al.]{lasby2023srigl}
M.~Lasby, A.~Golubeva, U.~Evci, M.~Nica, and Y.~Ioannou.
\newblock Dynamic sparse training with structured sparsity.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2024.

\bibitem[Mishra et~al.(2021)Mishra et~al.]{mishra2021accelerating}
Asit Mishra, Jorge Albericio Latorre, Jeff Pool, Darko Stosic, Dusan Stosic, Ganesh Venkatesh, Chong Yu, and Paulius Micikevicius.
\newblock Accelerating sparse deep neural networks.
\newblock {\em arXiv preprint arXiv:2104.08378}, 2021.

\bibitem[NVIDIA(2026)NVIDIA]{nvidiacusparselt}
NVIDIA.
\newblock cuSPARSELt: A high-performance CUDA library for sparse matrix-matrix multiplication.
\newblock \url{https://docs.nvidia.com/cuda/cusparselt/}, accessed 2026.

\bibitem[PyTorch(2026)PyTorch]{pytorchsemistructured}
PyTorch.
\newblock Semi-structured sparsity and \texttt{to\_sparse\_semi\_structured}.
\newblock \url{https://docs.pytorch.org/docs/2.12/sparse.html}, accessed 2026.

\bibitem[Mocanu et~al.(2018)Mocanu et~al.]{mocanu2018scalable}
D.~C. Mocanu, E.~Mocanu, P.~Stone, P.~H. Nguyen, M.~Gibescu, and A.~Liotta.
\newblock Scalable training of artificial neural networks with adaptive sparse connectivity inspired by network science.
\newblock {\em Nature Communications}, 9(1):2383, 2018.


\bibitem[ISIC(2024)ISIC]{isic2024slice3d}
International Skin Imaging Collaboration.
\newblock SLICE-3D 2024 Challenge Dataset.
\newblock International Skin Imaging Collaboration, 2024.
\newblock doi:10.34970/2024-slice-3d.
\newblock \url{https://doi.org/10.34970/2024-slice-3d}.

\bibitem[Micikevicius et~al.(2018)Micikevicius et~al.]{micikevicius2018mixed}
P.~Micikevicius, S.~Narang, J.~Alben, G.~Diamos, E.~Elsen, D.~Garcia, B.~Ginsburg, M.~Houston, O.~Kuchaiev, G.~Venkatesh, and H.~Wu.
\newblock Mixed precision training.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2018.

\bibitem[Charpentier et~al.(2020)Charpentier et~al.]{charpentier2020posterior}
B.~Charpentier, D.~Z{\"u}gner, and S.~G{\"u}nnemann.
\newblock Posterior Network: Uncertainty estimation without OOD samples via density based pseudo counts.
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)}, 2020.


\bibitem[Li et~al.(2022)Li et~al.]{li2022tlc}
B.~Li, Z.~Han, H.~Li, H.~Fu, and C.~Zhang.
\newblock Trustworthy long-tailed classification.
\newblock In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}, pages 6970--6979, 2022.

\bibitem[Sapkota and Yu(2023)Sapkota and Yu]{sapkota2023areo}
H.~Sapkota and Q.~Yu.
\newblock Adaptive robust evidential optimization for open set detection from imbalanced data.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2023.

\bibitem[Louizos et~al.(2017)Louizos et~al.]{louizos2017bayesian}
C.~Louizos, K.~Ullrich, and M.~Welling.
\newblock Bayesian Compression for Deep Learning.
\newblock In {\em Advances in Neural Information Processing Systems (NeurIPS)}, 2017.

\bibitem[Molchanov et~al.(2017)Molchanov et~al.]{molchanov2017variational}
D.~Molchanov, A.~Ashukha, and D.~Vetrov.
\newblock Variational Dropout and the Local Reparameterization Trick.
\newblock In {\em International Conference on Machine Learning (ICML)}, 2017.

\bibitem[Louizos et~al.(2018)Louizos et~al.]{louizos2018learning}
C.~Louizos, M.~Welling, and D.~P. Kingma.
\newblock Learning Sparse Neural Networks through $L_0$ Regularization.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2018.

\bibitem[Adamczewski and Park(2021)Adamczewski and Park]{adamczewski2021dirichlet}
K.~Adamczewski and M.~Park.
\newblock Dirichlet pruning for convolutional neural networks.
\newblock In {\em International Conference on Artificial Intelligence and Statistics (AISTATS)}, PMLR 130:3637--3645, 2021.

\bibitem[Ji et~al.(2024)Ji et~al.]{ji2024bidst}
J.~Ji, G.~Li, L.~Yin, M.~Qin, G.~Yuan, L.~Guo, S.~Liu, and X.~Ma.
\newblock Advancing dynamic sparse training by exploring optimization opportunities.
\newblock In {\em International Conference on Machine Learning (ICML)}, PMLR 235:21606--21619, 2024.

\bibitem[Tyagi et~al.(2026)Tyagi et~al.]{tyagi2025padst}
A.~Tyagi, A.~Iyer, L.~Young, W.~H. Renninger, C.~Kanan, and Y.~Zhu.
\newblock Efficient dynamic structured sparse training with learned shuffles.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2026.


\bibitem[Lin et~al.(2017)Lin et~al.]{lin2017focal}
T.-Y. Lin, P.~Goyal, R.~Girshick, K.~He, and P.~Doll{'a}r.
\newblock Focal loss for dense object detection.
\newblock In {\em Proceedings of the IEEE international conference on computer vision (ICCV)}, 2017.

\bibitem[J\o{}sang(2016)J\o{}sang]{josang2016subjective}
A.~J{\o}sang.
\newblock {\em Subjective Logic: A Formalism for Reasoning Under Uncertainty}.
\newblock Springer, 2016.

\bibitem[He et~al.(2016)He et~al.]{he2016deep}
K.~He, X.~Zhang, S.~Ren, and J.~Sun.
\newblock Deep residual learning for image recognition.
\newblock In {\em Proceedings of the IEEE conference on computer vision and pattern recognition (CVPR)}, 2016.

\bibitem[Loshchilov and Hutter(2017)Loshchilov and Hutter]{loshchilov2016sgdr}
I.~Loshchilov and F.~Hutter.
\newblock SGDR: Stochastic gradient descent with warm restarts.
\newblock In {\em International Conference on Learning Representations (ICLR)}, 2017.

\bibitem[Russakovsky et~al.(2015)Russakovsky et~al.]{russakovsky2015imagenet}
O.~Russakovsky, J.~Deng, H.~Su, J.~Krause, S.~Satheesh, S.~Ma, Z.~Huang, A.~Karpathy, A.~Khosla, M.~Bernstein, A.~C. Berg, and L.~Fei-Fei.
\newblock ImageNet Large Scale Visual Recognition Challenge.
\newblock {\em International Journal of Computer Vision (IJCV)}, 2015.

\end{thebibliography}

\ifwithappendix
\appendix

% Dense appendix diagnostics benefit from compact float spacing.  These
% settings apply only after the main paper and preserve normal float behavior.
\setcounter{topnumber}{4}
\setcounter{bottomnumber}{2}
\renewcommand{\topfraction}{0.95}
\renewcommand{\bottomfraction}{0.80}
\renewcommand{\textfraction}{0.05}
\renewcommand{\floatpagefraction}{0.75}
\setlength{\textfloatsep}{4pt plus 1pt minus 1pt}
\setlength{\floatsep}{3pt plus 1pt minus 1pt}
\setlength{\intextsep}{4pt plus 1pt minus 1pt}

\section*{Appendix Roadmap}
The appendix follows the order most useful for reproducibility and review.
Appendix~A gives the experimental setting, split rules, metric definitions,
hyperparameters, and release plan. Appendix~B provides the training loop and implementation
configuration. Appendix~C reports seed level ISIC results and sparse morphology
checks. Appendix~D contains component attribution diagnostics. Appendix~E gives
well posedness and algorithmic validity checks. Appendix~F clarifies the novelty
boundary against adjacent methods. Appendix~G reports external-domain,
softmax, and CIFAR scope analyses together with the hardware evidence taxonomy.

\section{Reproducibility Details}
\label{app:reproducibility}

\subsection*{1. Experimental Setting and Hyperparameters}
\begin{itemize}
    \item \textbf{Random Seeds:} Main ISIC comparisons use model seeds 42, 123, and 456 with split seed 42. Paper level component ablations include only variants that completed all three model seeds. Partial single seed or two seed ablations are excluded from the reported ablation table.
    \item \textbf{Preprocessing \& Dimensions:} Images are resized to $224 \times 224$ and normalized using ImageNet channel wise statistics: $\mu = [0.485, 0.456, 0.406]$ and $\sigma = [0.229, 0.224, 0.225]$. Data augmentation includes horizontal and vertical flips.
    \item \textbf{Optimizer and Learning Rate Schedule:} We employ the AdamW optimizer ($\beta_1=0.9$, $\beta_2=0.999$, weight decay $10^{-4}$ on active weights). The learning rate is initialized at $1\times10^{-6}$ and linearly warmed up to $4\times10^{-5}$ over the first epoch, followed by a cosine annealing schedule \cite{loshchilov2016sgdr} over 40 epochs.
\end{itemize}

\subsection*{2. Evaluation Protocols}
\begin{itemize}
    \item \textbf{Validation/Test Split Rules:} The ISIC 2024 dataset is partitioned into 70\% training, 5\% validation, 5\% calibration, and 20\% test sets using split seed 42. The resulting split contains 5,775 training images, 21,930 validation images, 24,235 calibration images, and 65,931 test images, with 77 malignant and 65,854 benign test cases. Patient-grouped stratification prevents a patient from appearing in more than one partition. The test set is excluded from training, calibration, model selection, and threshold selection and is used only for final reporting after each configuration is fixed.
    \item \textbf{Threshold Selection Protocol:} For ISIC, operating points are selected by performing threshold sweeps on the isolated validation partition to hit target clinical sensitivities ($\ge 80\%$). Temperature and bias parameters are fitted on the separate calibration partition. Validation derived thresholds are frozen before final test evaluation.
    \item \textbf{ECE Binning and pAUC:} The reported ECE is adaptive ECE with 15 equal frequency bins. Equal width ECE is also computed as an auxiliary diagnostic. pAUC$_{0.80}$ is the unnormalized partial area under the ROC curve within $\text{TPR} \in [0.80, 1.0]$ (maximum value $0.20$), computed with trapezoidal interpolation and arithmetically averaged across seeds.
\end{itemize}

\subsection*{3. Metric Definitions and Interpretation}\label{app:metrics}
The evaluation separates operating-point metrics, threshold-independent ranking,
calibration, selective prediction, external-domain separation, and systems
measurements. Values from different families answer different questions and
should not be compared numerically as if they shared one scale.

\paragraph{Operating-point classification metrics.}
For malignant class $1$, sensitivity and specificity are
\begin{equation}\label{eq:appendix_operating_metrics}
\begin{aligned}
\mathrm{TPR}&=\frac{\mathrm{TP}}{\mathrm{TP}+\mathrm{FN}},\\
\mathrm{TNR}&=\frac{\mathrm{TN}}{\mathrm{TN}+\mathrm{FP}},\\
\mathrm{BalAcc}&=\frac{\mathrm{TPR}+\mathrm{TNR}}{2}.
\end{aligned}
\end{equation}
Sensitivity measures the fraction of malignant images retrieved; specificity
measures the fraction of benign images rejected. Balanced accuracy weights the
two classes equally and therefore avoids the majority-class dominance of raw
accuracy. All three depend on a decision threshold. The implementation reports
the default threshold $0.5$, a validation-selected balanced-accuracy threshold,
and the clinical threshold in Eq.~\ref{eq:threshold}. Thresholds are never
selected on the test set. Consequently, a balanced-accuracy value cannot be
interpreted without its associated threshold and sensitivity--specificity pair.

\paragraph{High-sensitivity and global ranking metrics.}
Let $q=1-\mathrm{TPR}$ denote the false-negative rate and let
$\mathrm{TNR}(q)$ be specificity along the ROC curve. The official-style
unnormalized high-sensitivity partial area is
\begin{equation}\label{eq:appendix_ranking_metrics}
\begin{aligned}
\mathrm{pAUC}_{0.80}&=\int_{0}^{0.20}\mathrm{TNR}(q)\,dq,\\
\mathrm{M\text{-}AUROC}&=\frac{1}{K}\sum_{c=1}^{K}
\int_{0}^{1}\mathrm{TPR}_c(r)\,dr,\\
r&=\mathrm{FPR}_c.
\end{aligned}
\end{equation}
pAUC$_{0.80}$ evaluates specificity only where sensitivity is at least $0.80$;
its range is $[0,0.20]$, so $0.1172$ is an area on a $0.20$ scale rather than
$11.72\%$ classification accuracy. Macro AUROC averages one-vs-rest ROC areas
and summarizes ranking over all thresholds. In the binary implementation it is
computed from the malignant probability through \texttt{roc\_auc\_score}; the
``Macro'' name is retained for consistency with the multiclass evaluation
interface. AUROC is useful for global discrimination but does not determine a
clinically acceptable operating threshold.

Average precision is
\begin{equation}\label{eq:appendix_ap}
\mathrm{AP}=\sum_{k}(R_k-R_{k-1})P_k,
\end{equation}
where $(P_k,R_k)$ are precision and recall at successive ranked-score cutoffs.
The implementation uses \texttt{average\_precision\_score}; therefore the paper
uses ``AP'' rather than trapezoidal ``PR AUC.'' AP emphasizes retrieval quality
for the rare malignant class and is sensitive to positive prevalence. A method
can consequently improve pAUC$_{0.80}$ while reducing AP if it ranks enough
malignant cases highly for sensitivity but introduces more false positives near
the top of the list.

\paragraph{Calibration and selective prediction.}
For each image, confidence is $c_n=\max_c\hat p_{n,c}$ and correctness is
$a_n=\mathbb 1[\hat y_n=y_n]$ under the default decision. After sorting by
confidence, adaptive ECE divides the test set into 15 approximately
equal-frequency bins $B_b$:
\begin{equation}\label{eq:appendix_ece}
\begin{aligned}
\mathrm{ECE}&=\sum_{b=1}^{15}\frac{|B_b|}{N}
\left|\mathrm{acc}(B_b)-\mathrm{conf}(B_b)\right|,\\
\mathrm{acc}(B_b)&=\frac{1}{|B_b|}\sum_{n\in B_b}a_n,\\
\mathrm{conf}(B_b)&=\frac{1}{|B_b|}\sum_{n\in B_b}c_n.
\end{aligned}
\end{equation}
Lower ECE indicates closer agreement between confidence and empirical
correctness; it does not distinguish overconfidence from underconfidence unless
the signed bin gaps or reliability diagram are also inspected. Under extreme
negative prevalence, a classifier that is consistently confident on benign
images can obtain a small aggregate ECE despite weak malignant retrieval.
Class-conditional ECE is therefore retained as an auxiliary diagnostic.

For selective prediction, samples are sorted by decreasing confidence. At
coverage $m/N$, risk is the cumulative default-decision error among the first
$m$ retained samples:
\begin{equation}\label{eq:appendix_aurc}
\begin{aligned}
\mathrm{Risk}(m/N)&=\frac{1}{m}\sum_{i=1}^{m}
\mathbb 1[\hat y_{(i)}\ne y_{(i)}],\\
\mathrm{AURC}&=\int_0^1\mathrm{Risk}(c)\,dc.
\end{aligned}
\end{equation}
Lower AURC means errors are concentrated among low-confidence cases that can be
deferred earlier. AURC depends on both base classification error and confidence
ordering, so it should not be read as a pure calibration metric.

\paragraph{Patient-grouped top-15 sensitivity and external OOD metrics.}
For every patient containing at least one malignant lesion, images are ranked by
malignant probability and only the top 15 are retained. The reported statistic
is
\begin{equation}\label{eq:appendix_top15}
\mathrm{SE}_{\mathrm{top\text{-}15}}
=\frac{\displaystyle\sum_{p}\sum_{i\in\operatorname{Top15}(p)}y_{p,i}}
{\displaystyle\sum_{p}\sum_{i\in p}y_{p,i}}.
\end{equation}
Thus it is lesion-level malignant recall after patient-wise top-15 triage, not
the fraction of patients with at least one detected malignancy. Higher values
mean that a fixed review budget retains more malignant lesions.

For the PAD-UFES-20 diagnostic, ISIC images receive OOD label $0$ and external
images receive label $1$. OOD AUROC measures pairwise ordering of the fixed
layer-3 KNN score, while OOD AUPR is average precision with external images as
positives. Both are threshold-independent and higher is better. They measure
source-to-external feature separation, not lesion classification, calibration,
fairness, or superiority of Dirichlet uncertainty.

\paragraph{Sparsity, computation, and runtime metrics.}
The structural and backend quantities are defined as
\begin{equation}\label{eq:appendix_hardware_metrics}
\begin{aligned}
\mathrm{Density}&=\frac{N_{\mathrm{active}}}{N_{\mathrm{eligible}}},\\
\mathrm{Valid}_{2:4}&=\frac{N_{\mathrm{valid\ 2:4\ blocks}}}
{N_{\mathrm{eligible\ blocks}}},\\
\mathrm{MACRed.}&=1-\frac{\mathrm{GMAC}_{\mathrm{active}}}
{\mathrm{GMAC}_{\mathrm{dense}}},\\
\mathrm{Speedup}&=\frac{\mathrm{images/s}_{\mathrm{sparse}}}
{\mathrm{images/s}_{\mathrm{reference}}}.
\end{aligned}
\end{equation}
Wrapped-layer density is exactly $50\%$ under 2:4, whereas whole-model density
also includes the dense stem and head. Active coordinates and theoretical MACs
describe executable arithmetic, not serialized file size. Throughput is images
processed per second; latency is GPU compute time per batch. A same-graph sparse
toggle isolates sparse-tactic execution more directly than comparing separately
built networks. All runtime ratios remain conditional on device, batch size,
precision, graph, and TensorRT tactic selection; a batch-64 speedup does not
imply a batch-1 latency gain.

\subsection*{4. Source Code and Data Releases}
\begin{itemize}
    \item \textbf{Submission Artifact:} The anonymized supplementary archive accompanying submission contains the sparse wrapper layers, loss definitions, experiment runners, environment requirements, fixed seeds, and evaluation scripts needed to reconstruct the reported protocol. The artifact manifest records the source revision; no post-publication release promise is used as evidence of reproducibility.
\end{itemize}

\section{DST-EDL Algorithm and Implementation Details}
\label{app:algorithm}
This appendix gives the procedural details that are referenced from the
optimization section. It contains the warmup rule, the post hoc evidential
calibration formula, the full training loop, and the implementation
configuration used for the reported ISIC 2024 case study.

\paragraph{Learning rate and loss scale warmup.}
During the first epoch, the learning rate is linearly warmed from $\eta_{\text{init}}=10^{-6}$ to $\eta_{\text{base}}=4\times10^{-5}$. An initial loss multiplier decays from $S_{\max}=4.0$ to $1.0$:
\begin{equation}\label{eq:warmup_main}
\begin{aligned}
    \eta_s&=\eta_{\text{init}}
    +(\eta_{\text{base}}-\eta_{\text{init}})
    \frac{s}{S_{\text{warmup}}},\\
    S_{\text{loss}}(s)&=S_{\max}
    -(S_{\max}-1)\frac{s}{S_{\text{warmup}}}.
\end{aligned}
\end{equation}
The scaled objective $L_{\text{scaled}}=S_{\text{loss}}(s)L_{\text{total}}$ is used during warmup and $L_{\text{total}}$ afterward.

\paragraph{Bias corrected evidential calibration.}
After training, logits are corrected before the Softplus evidence map by
\begin{equation}
\begin{aligned}
    z^{\mathrm{cal}}_c&=
    \frac{z^{\mathrm{base}}_c+\Delta^\pi_c}{T_{\mathrm{cal}}}
    +b_{\mathrm{res},c},\\
    \Delta^\pi_c&=\log(\pi_{\mathrm{true},c}+\epsilon)\\
    &\quad-\log(\pi_{\mathrm{train},c}+\epsilon).
\end{aligned}
\end{equation}
Temperature only calibration sets $\bm{b}_{\mathrm{res}}=\bm{0}$. The reported bias temperature mode learns both $T_{\mathrm{cal}}$ and $\bm{b}_{\mathrm{res}}$ on the calibration partition.

\begin{table}[tbp]
\centering
\caption{DST-EDL training and optimization loop.}
\label{alg:dst_edl}
\scriptsize
\renewcommand{\arraystretch}{1.04}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}r p{0.82\columnwidth}@{}}
\toprule
1 & Initialize data $\mathcal{D}$, weights $\bm{W}^{(0)}$, latent scores $\bm{V}^{(0)}$, epochs $T$, and warmup $T_w$. \\
2 & For each epoch $t=1,\ldots,T$, keep the reported identity channel maps fixed; at $t=T_w$, initialize latent scores from warmup weights. \\
3 & If $t\ge T_w$, concatenate the first four shuffled training mini-batches (at most 128 images); use this natural-frequency proxy for all structural signals. \\
4 & Compute $R_{\mathcal B}$ and $L_{\text{grow}}$, then store pruner and regrower gradients. \\
5 & Compute $C_{ij}$ by Eq.~\ref{eq:C_ij} and $G_{ij}$ by Eq.~\ref{eq:G_ij}. \\
6 & If an entire layer has flat regrowth rank, add the vacuity-ranked tie-breaker $\bm\Xi^{(l)}$ to dormant coordinates. \\
7 & Update latent scores $\bm{V}$ by Eq.~\ref{eq:score_update}. \\
8 & Cache 2:4 masks $\bm{M}^{(l)}$ using Eq.~\ref{eq:mask}. During warmup set masks to dense. \\
9 & For each mini batch, compute $\eta_s$ and $S_{\text{loss}}(s)$ by Eq.~\ref{eq:warmup_main}. \\
10 & Forward with $\bm{W}_{\text{eff}}=\bm{W}\odot\bm{M}$, set $\alpha_c=\ln(1+e^{z_c})+1$, and compute uncertainties. \\
11 & Compute $L_{\text{scaled}}=S_{\text{loss}}(s)L_{\text{EFL}}$ using Eq.~\ref{eq:efl} and update active weights. \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\paragraph{Layerwise tie breaking.}
The implementation gates anti-crystallization at layer level rather than per coordinate. For layer $l$,
\begin{equation}\label{eq:anticryst_exact}
\begin{aligned}
\bm\Xi^{(l)}&=
\mathbb 1\!\left[\max_{ij}G_{ij}^{(l)}\le10^{-8}\right]\\
&\quad\times\left[0.0316\,\bm\varepsilon^{(l)}
\odot\bm Q^{(l)}\right]_+,\\
\bm\varepsilon^{(l)}&\sim\mathcal N(\bm0,\bm I).
\end{aligned}
\end{equation}
where $\bm Q^{(l)}$ broadcasts the percentile rank of node-level vacuity across the layer's weight coordinates. This nonnegative term is added only to the dormant branch of Eq.~\ref{eq:score_step}; it is zero whenever KL-based regrowth provides a nonflat ordering.

\begin{table}[tbp]
\centering
\caption{Training and calibration configuration for the ISIC 2024 case study.}
\label{tab:config}
\scriptsize
\renewcommand{\arraystretch}{0.96}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}p{0.30\columnwidth} p{0.24\columnwidth} p{0.36\columnwidth}@{}}
\toprule
\textbf{Component} & \textbf{Value} & \textbf{Role} \\
\midrule
Backbone & ResNet-18 \cite{he2016deep} & ImageNet-1K pretrained \cite{russakovsky2015imagenet} \\
Classes ($K$) & 2 & Benign / malignant \\
Dense warmup mask & $\bm{1}$ & No pruning before $T_w$ \\
Warmup EFL exponent & $0.0$ & Effective focal exponent during dense warmup \\
2:4 mask & top two of four & Project latent scores $\bm{V}$ \\
Topology refresh & Cached once/epoch & Up to 4 train mini batches \\
Post hoc calibration & Bias temperature & Fit $T_{\text{cal}}$ and residual bias \\
Optimizer & AdamW & $\beta_1=0.9$, $\beta_2=0.999$ \\
Learning rate & $4{\times}10^{-5}$ & Linear warmup from $10^{-6}$ \\
Weight decay & $10^{-4}$ & Active weights \\
Gradient clipping & $1.0$ & Global $L_2$ norm \\
Batch size / epochs & 32 / 40 & ISIC repeated seed runs \\
Dense topology warmup & 12 epochs & 30\% of training budget \\
EFL $\gamma_{\text{base}}$ & $5.0$ & Cosine decay after warmup \\
$\lambda_{\text{KL}}$, $t_{\text{anneal}}$ & $0.1$, 10 & KL strength and annealing \\
$\omega_{\max}$ & $10.0$ & Cap class weights \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\begin{table}[tbp]
\centering
\caption{Structural controller hyperparameters for reported DST-EDL runs.}
\label{tab:struct_config}
\scriptsize
\renewcommand{\arraystretch}{0.98}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}p{0.34\columnwidth} p{0.23\columnwidth} p{0.35\columnwidth}@{}}
\toprule
\textbf{Parameter} & \textbf{Value} & \textbf{Role} \\
\midrule
$\beta_m$ & $0.95$ & Structural momentum \\
$\eta_{\text{struct}}$ & $0.02$ & Latent score step size \\
$\rho$ & $0.25$ & Dormant link attenuation \\
$V_{\max}$ & $5.0$ & Latent score clamp \\
Pruner & Soft signed first order & Eq.~\ref{eq:C_ij}, strength $0.5$ \\
Regrower & KL uniform & Eq.~\ref{eq:grow_objective} \\
Exploration scale & $0.0316$ & Anti crystallization near flat scores \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\section{Seed Level Results and Sparse Morphology Diagnostics}
\label{app:seed_level}
This appendix provides the per seed evidence behind the aggregated ISIC 2024
numbers in the main paper, followed by the sparse morphology checks used to
verify exact valid block 2:4 feasibility. The tables report operating point
metrics, ranking metrics, structural validity, and hardware cost diagnostics in
the same order as the main results discussion.

\begin{table}[tbp]
\centering
\caption{Seed level DST-EDL operating points on ISIC 2024 for split seed 42. Lower ECE is better.}
\label{tab:seed_level}
\scriptsize
\renewcommand{\arraystretch}{1.05}
\setlength{\tabcolsep}{2.0pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l C C C C@{}}
\toprule
\textbf{Seed} & \textbf{Bal.} $\uparrow$ & \textbf{Sens.} $\uparrow$ & \textbf{Spec.} $\uparrow$ & \textbf{ECE} $\downarrow$ \\
\midrule
42 & $0.7976$ & $0.6623$ & $0.9329$ & $0.0560$ \\
123 & $0.7998$ & $0.7922$ & $0.8073$ & $0.0547$ \\
456 & $0.7358$ & $0.5065$ & $0.9651$ & $0.0555$ \\
\midrule
Mean $\pm$ Std & $0.7777\pm0.0363$ & $0.6537\pm0.1430$ & $0.9018\pm0.0834$ & $0.0554\pm0.0007$ \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\begin{table}[tbp]
\centering
\caption{Seed level DST-EDL ranking metrics on ISIC 2024. SE$_{\mathrm{top\mbox{-}15}}$ is patient-level sensitivity among the top 15 ranked cases.}
\label{tab:seed_level_ranking}
\scriptsize
\renewcommand{\arraystretch}{1.05}
\setlength{\tabcolsep}{2.0pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l C C C C@{}}
\toprule
\textbf{Seed} & \textbf{pAUC$_{0.80}$} $\uparrow$ & \textbf{Macro AUROC} $\uparrow$ & \textbf{AP} $\uparrow$ & \textbf{SE$_{\mathrm{top\mbox{-}15}}$} $\uparrow$ \\
\midrule
42 & $0.1211$ & $0.8913$ & $0.0371$ & $0.5195$ \\
123 & $0.1095$ & $0.8751$ & $0.0336$ & $0.4935$ \\
456 & $0.1210$ & $0.8894$ & $0.0298$ & $0.5065$ \\
\midrule
Mean $\pm$ Std & $0.1172\pm0.0067$ & $0.8853\pm0.0089$ & $0.0335\pm0.0037$ & $0.5065\pm0.0130$ \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

Table~\ref{tab:structural_diagnostics} reports the corresponding structural
validity diagnostics for the sparse models.

\begin{table}[tbp]
\centering
\caption{Structural diagnostics on ISIC 2024. Sparse models maintained 50.0\% active density with no detected structural collapse. The current DST-EDL runs satisfy all complete valid 2:4 blocks in the wrapped sparse layers.}
\label{tab:structural_diagnostics}
\footnotesize
\renewcommand{\arraystretch}{1.1}
\setlength{\tabcolsep}{2pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l c c c@{}}
\toprule
\textbf{Model} & \textbf{Density} & \textbf{Valid 2:4} & \textbf{Collapse} \\
\midrule
Static 2:4 EDL & 50.0\% & Exact 2:4 & No \\
RigL-style 2:4 & 50.0\% & Exact 2:4 & No \\
\textbf{DST-EDL} & 50.0\% & Exact 2:4 & No \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\begin{table}[tbp]
\centering
\caption{Hardware cost diagnostic for ResNet 18 at image size $224\times224$ and batch size 64. MAC and FLOP values are reported per image. Throughput is measured with masked PyTorch execution. Masked PyTorch does not realize sparse-kernel acceleration; however, a separate TensorRT diagnostic shows backend-specific sparse-path speedup under the tested configuration.}
\label{tab:structural_cost}
\scriptsize
\renewcommand{\arraystretch}{1.05}
\setlength{\tabcolsep}{1.2pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l c c c c c@{}}
\toprule
\textbf{Model} & \textbf{Active} & \textbf{Valid} & \textbf{GMAC} & \textbf{MAC red.} & \textbf{Img/s} \\
\midrule
Dense EDL & 100\% & N/A & 1.8136 & 0.0\% & 1358.1 \\
Static 2:4 EDL & 50.09\% & Exact & 0.9658 & 46.75\% & 622.6 \\
\textbf{DST-EDL} & 50.09\% & Exact & 0.9658 & 46.75\% & 617.2 \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

The sparse rows reduce theoretical active cost from 1.8136 to 0.9658 GMAC per
image. This corresponds to a 46.75\% total model MAC reduction and a 50.0\%
wrapped layer MAC reduction. The same accounting gives 3.6271 dense GFLOPs per
image and 1.9316 theoretical active GFLOPs per image for the sparse rows. The
measured throughput in Table~\ref{tab:structural_cost} is lower than dense because the implementation uses masked
PyTorch kernels without a compressed 2:4 sparse kernel. 

The fixed predictive-evaluation checkpoints are not reused for backend acceleration evaluation. We separately benchmark ONNX exports from the hardware-compatible deployment configuration with TensorRT 11.1 on an RTX A2000 8GB Laptop GPU. Runs use batch 64, FP16 bindings, disabled host--device transfers, CUDA Graph execution, and repeated timing after warmup. In the controlled comparison, the graph is frozen and only TensorRT sparse execution is toggled: throughput changes from 4847.1 to 5546.3 images/s and mean GPU latency from 13.20 to 11.54 ms, a 1.14$\times$ throughput gain and 12.6\% latency reduction. Relative to Dense EDL in the same TensorRT setting, the hardware-compatible configuration is 1.18$\times$ faster (Table~\ref{tab:hardware_benchmarks}). Build logs show sparse tactics in 10/19 eligible DST-EDL layers, versus 12/19 for Static 2:4 and 13/19 for RigL-style 2:4. These measurements establish sparse-path execution for this graph and device; without predictive metrics for the exported checkpoint, they do not establish an end-to-end accuracy--efficiency tradeoff.

\begin{table}[!t]
\centering
\caption{TensorRT FP16 deployment diagnostic on RTX A2000 (TensorRT 11.1, batch 64, CUDA Graph, H2D disabled). The first block compares deployment configurations; the second toggles sparse execution in one frozen DST-EDL graph.}
\label{tab:hardware_benchmarks}
\scriptsize
\renewcommand{\arraystretch}{1.05}
\setlength{\tabcolsep}{1.2pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l c c c c c@{}}
\toprule
\textbf{Model / comparison} & \textbf{TRT path} & \textbf{Sparse tactics} & \textbf{Img/s} $\uparrow$ & \textbf{Latency} $\downarrow$ & \textbf{Speedup} \\
\midrule
Dense EDL & disable & N/A & 4686.9 & 13.65 ms & 1.00$\times$ \\
Static 2:4 EDL & enable & 12/19 & 5699.8 & 11.23 ms & 1.22$\times$ \\
RigL-style 2:4 & enable & 13/19 & 5593.7 & 11.44 ms & 1.19$\times$ \\
\textbf{DST-EDL} & enable & 10/19 & \textbf{5546.3} & \textbf{11.54 ms} & \textbf{1.18$\times$} \\
\midrule
DST-EDL, same graph & disable & N/A & 4847.1 & 13.20 ms & 1.00$\times$ \\
\textbf{DST-EDL, same graph} & enable & 10/19 & \textbf{5546.3} & \textbf{11.54 ms} & \textbf{1.14$\times$} \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\FloatBarrier
\section{Component Attribution Diagnostics}
\label{app:ablation}
This appendix reports diagnostic component variants used for attribution.
The variants are component probes around the reported method, not separate full
models. Table~\ref{tab:ablation_components_full} includes only variants that
completed all three runs with seeds 42, 123, and 456. Variants without all
three runs are omitted from the paper-level ablation table.

\begin{table}[!t]
\centering
\caption{Three seed component operating point diagnostics on ISIC 2024. Only variants with all seeds 42, 123, and 456 are included. Lower ECE is better.}
\label{tab:ablation_components_full}
\scriptsize
\renewcommand{\arraystretch}{1.05}
\setlength{\tabcolsep}{1.8pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l C C C C@{}}
\toprule
\textbf{Variant} & \textbf{Bal.} $\uparrow$ & \textbf{Sens.} $\uparrow$ & \textbf{Spec.} $\uparrow$ & \textbf{ECE} $\downarrow$ \\
\midrule
\textbf{DST-EDL} & $\mathbf{0.8110\pm0.0050}$ & $\mathbf{0.7835\pm0.0418}$ & $0.8386\pm0.0399$ & $0.0564\pm0.0009$ \\
w/o pruner & $0.7946\pm0.0077$ & $0.7229\pm0.0492$ & $0.8662\pm0.0394$ & $0.0552\pm0.0007$ \\
w/o regrower & $0.7046\pm0.0527$ & $0.6753\pm0.0723$ & $0.7339\pm0.0520$ & $0.0556\pm0.0027$ \\
asymmetric KL & $0.8033\pm0.0116$ & $0.7229\pm0.0715$ & $0.8836\pm0.0485$ & $0.0556\pm0.0003$ \\
w/o EFL & $0.7500\pm0.0266$ & $0.5628\pm0.1050$ & $\mathbf{0.9372\pm0.0519}$ & $\mathbf{0.0544\pm0.0003}$ \\
w/o anti cryst. & $0.7837\pm0.0369$ & $0.6753\pm0.1498$ & $0.8920\pm0.0761$ & $0.0556\pm0.0006$ \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\begin{table}[!t]
\centering
\caption{Three-seed component ranking and risk diagnostics on ISIC 2024. Bold marks the largest observed mean, not significance. Lower AURC is better.}
\label{tab:ablation_ranking}
\scriptsize
\renewcommand{\arraystretch}{1.05}
\setlength{\tabcolsep}{1.8pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l C C C C@{}}
\toprule
\textbf{Variant} & \textbf{pAUC$_{0.80}$} $\uparrow$ & \textbf{Macro AUROC} $\uparrow$ & \textbf{AP} $\uparrow$ & \textbf{AURC} $\downarrow$ \\
\midrule
\textbf{DST-EDL} & $0.1168\pm0.0070$ & $\mathbf{0.8891\pm0.0080}$ & $\mathbf{0.0355\pm0.0063}$ & $\mathbf{0.00023\pm0.00006}$ \\
w/o pruner & $0.1051\pm0.0057$ & $0.8720\pm0.0069$ & $0.0319\pm0.0029$ & $0.00027\pm0.00006$ \\
w/o regrower & $0.0640\pm0.0187$ & $0.7663\pm0.0629$ & $0.0077\pm0.0049$ & $0.00057\pm0.00021$ \\
asymmetric KL & $0.1127\pm0.0141$ & $0.8845\pm0.0159$ & $0.0348\pm0.0042$ & $0.00033\pm0.00006$ \\
w/o EFL & $0.0898\pm0.0010$ & $0.8546\pm0.0016$ & $0.0298\pm0.0038$ & $0.00030\pm0.00000$ \\
w/o anti cryst. & $\mathbf{0.1173\pm0.0098}$ & $0.8857\pm0.0109$ & $0.0352\pm0.0047$ & $0.00030\pm0.00010$ \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

Within the implemented DST-EDL system, removing regrowth produces the largest
observed degradation: pAUC$_{0.80}$ changes from $0.1168$ to $0.0640$, Macro
AUROC from $0.8891$ to $0.7663$, AP from $0.0355$ to $0.0077$, and AURC
from $0.00023$ to $0.00057$. This supports the need for dormant-link
reactivation in this controller, but the ablation does not compare KL regrowth
against every alternative regrowth rule. Removing the pruner yields smaller
decreases in balanced accuracy, sensitivity, pAUC$_{0.80}$, and Macro AUROC.
Asymmetric KL remains close in ranking metrics but moves the operating point
toward higher specificity and lower sensitivity. Removing focal modulation
improves ECE and specificity while sensitivity falls from $0.7835$ to $0.5628$.
Removing anti-crystallization gives a slightly higher mean pAUC$_{0.80}$
($0.1173$ vs. $0.1168$) and a larger sensitivity SD. These are descriptive
component effects across three runs; no significance claim is made.

\FloatBarrier
\section{Well Posedness and Algorithmic Validity Checks}
\label{app:proofs}

\begin{table}[!t]
\centering
\caption{Scope of formal claims made for DST-EDL.}
\label{tab:guarantees}
\scriptsize
\renewcommand{\arraystretch}{0.98}
\begin{tabular}{@{}p{0.47\columnwidth}p{0.43\columnwidth}@{}}
\toprule
\textbf{Established by design or proof} & \textbf{Not claimed} \\
\midrule
Positive Dirichlet parameters, bounded vacuity, and non negative ambiguity proxy &
Exact Bayesian epistemic correctness \\
Exact valid block 2:4 feasibility over eligible complete 4-blocks &
Global convergence or topology optimality \\
Bounded latent topology scores and deterministic top two projection &
Uniform superiority over dense or dynamic sparse baselines \\
Ordinary masked gradient isolation plus virtual probe structural saliency &
Validation improvement without empirical evidence \\
Local first order pruning/regrowth interpretations and proxy batch concentration &
Universal hardware speedup or batch-1 latency improvement \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Compact Dirichlet Identities}
For completeness, the Dirichlet density for $\bm{p}$ on the probability simplex is
\begin{equation}\label{eq:dirichlet_pdf_app}
\begin{aligned}
    \Dir(\bm p\mid\bm\alpha)
    &=\frac{1}{B(\bm\alpha)}
    \prod_{c=1}^{K}p_c^{\alpha_c-1}\\
    &=\frac{\Gamma(S)}
    {\prod_{c=1}^{K}\Gamma(\alpha_c)}
    \prod_{c=1}^{K}p_c^{\alpha_c-1}.
\end{aligned}
\end{equation}
The implementation uses the following identities from the main text:
\begin{equation}
\begin{aligned}
    \hat p_c&=\frac{\alpha_c}{S},\\
    u_e&=\frac{K}{S},\\
    u_a&=\sum_{c=1}^{K}\frac{\alpha_c}{S}
    \left[\psi(S+1)-\psi(\alpha_c+1)\right].
\end{aligned}
\end{equation}
Under the Softplus evidence parameterization, $\alpha_c>1$ for finite logits. Therefore $S>K$. Hence $u_e\in(0,1)$. Its supremum $1$ is approached as all evidence tends to zero. Also, $S\ge \alpha_c$ and $\psi(\cdot)$ is monotone on positive inputs. Each entropy summand is non negative. Hence $u_a\ge0$.

\subsection{Structural Risk Ratio Chain Rule}
\label{app:r_chain}
The batch structural risk ratio $R_{\mathcal B}$ in Eq.~\ref{eq:batch_risk} is differentiable for $\epsilon>0$ under the Softplus evidence map when the current mask is held fixed during the proxy structural pass. In this derivation, $w_{\text{eff},ij}$ abbreviates the corresponding coordinate of the virtual tensor $\widetilde{\bm W}_{\text{eff}}$; for an active link it has the same value as $W_{ij}$. Write $\partial_{ij}$ for differentiation with respect to that virtual coordinate. The derivative is the batch average of the per-sample ratio derivatives:
\begin{equation}
\begin{aligned}
    \partial_{ij}R_{\mathcal B}
    &=\frac{1}{|\mathcal B|}\sum_{n\in\mathcal B}
    \left[\frac{\partial_{ij}u_a}{u_e+\epsilon}\right.\\
    &\left.\quad-\frac{u_a\,\partial_{ij}u_e}
    {(u_e+\epsilon)^2}\right]^{(n)}.
\end{aligned}
\end{equation}
At the scalar level, $r(u_a,u_e)=u_a/(u_e+\epsilon)$ satisfies
\begin{equation*}
\begin{aligned}
    \frac{\partial r}{\partial u_a}&=\frac{1}{u_e+\epsilon}>0,\\
    \frac{\partial r}{\partial u_e}&=-\frac{u_a}{(u_e+\epsilon)^2}\le0.
\end{aligned}
\end{equation*}
The proxy is monotone increasing in ambiguity and monotone decreasing in vacuity whenever $u_a\ge0$.
The vacuity derivative expands through the Dirichlet parameters as
\begin{equation}
\begin{aligned}
    \frac{\partial u_e}{\partial w_{\mathrm{eff},ij}}
    &=\sum_{m=1}^{K}
    \frac{\partial u_e}{\partial\alpha_m}
    \frac{\partial\alpha_m}{\partial z_m}\\
    &\quad\times\frac{\partial z_m}{\partial w_{\mathrm{eff},ij}},\\
    \frac{\partial u_e}{\partial\alpha_m}&=-\frac{K}{S^2}.
\end{aligned}
\end{equation}
where $\frac{\partial \alpha_m}{\partial z_m}=1/(1+\exp(-z_m))$ for $\alpha_m=\ln(1+\exp(z_m))+1$. For the entropy proxy, let $\psi_1(\cdot)$ denote the trigamma function and $\delta_{cm}$ the Kronecker delta. Then
\begin{equation}
\begin{aligned}
    \frac{\partial u_a}{\partial\alpha_m}
    ={}&\sum_{c=1}^{K}\frac{\delta_{cm}S-\alpha_c}{S^2}\\
    &\quad\times\left[\psi(S+1)-\psi(\alpha_c+1)\right]\\
    &+\sum_{c=1}^{K}\frac{\alpha_c}{S}\\
    &\quad\times\left[\psi_1(S+1)-\delta_{cm}\psi_1(\alpha_c+1)\right].
\end{aligned}
\end{equation}
This gives
\begin{equation}
\begin{aligned}
    \frac{\partial u_a}{\partial w_{\mathrm{eff},ij}}
    &=\sum_{m=1}^{K}
    \frac{\partial u_a}{\partial\alpha_m}
    \frac{\partial\alpha_m}{\partial z_m}\\
    &\quad\times\frac{\partial z_m}{\partial w_{\mathrm{eff},ij}}.
\end{aligned}
\end{equation}
This expansion provides the virtual-coordinate derivative used by the signed pruning score in Eq.~\ref{eq:C_ij}. When the coordinate is active, $M_{ij}=1$ and $W_{ij}=\widetilde W_{\text{eff},ij}$ numerically, so the Taylor removal term is the one used in the main equation. The structural gradient is nevertheless detached and is not an optimizer update. The expression combines local vacuity and expected-entropy derivatives through the network Jacobian $\partial z_m/\partial w_{\text{eff},ij}$.

\subsection{Proxy Batch Concentration of the Risk Ratio Signal}
\label{app:proxy_concentration}
Fix the current weights and mask during a structural refresh, and let
$r(x)=u_a(x)/(u_e(x)+\epsilon)$ under the proxy sampling distribution.
For a $K$-class categorical variable, entropy is bounded above by
$\log K$, hence the Dirichlet expected categorical entropy satisfies
$0\le u_a(x)\le\log K$. Since $u_e(x)+\epsilon\ge\epsilon$, we have
\[
    0\le r(x)\le \frac{\log K}{\epsilon}.
\]
Let $\bar R=\mathbb{E}[r(x)]$ and
$R_{\mathcal B}=|\mathcal B|^{-1}\sum_{x\in\mathcal B}r(x)$ for an
i.i.d. proxy batch from this sampling distribution. Hoeffding's
inequality gives, for any $\delta\in(0,1)$, with probability at least
$1-\delta$,
\[
    |R_{\mathcal B}-\bar R|
    \le
    \frac{\log K}{\epsilon}
    \sqrt{\frac{\log(2/\delta)}{2|\mathcal B|}}.
\]
This result controls the estimation noise of the scalar proxy under the chosen
proxy sampling distribution. Validation risk improvement, convergence, and
dominance over magnitude or task gradient based sparse training remain outside
the statement.

\subsection{KL Regularization Note}
The reported DST-EDL configuration uses the symmetric KL penalty in Eq.~\ref{eq:kl_loss_main}: the KL multiplier is $1.0$ for every sample and class. The adjusted false evidence parameter $\tilde{\alpha}_c^{(n)}=y_c^{(n)}+(1-y_c^{(n)})\alpha_c^{(n)}$ defines the compact KL term. Its expanded form is the standard Dirichlet to Dirichlet KL with the true class concentration removed from the shrinkage penalty:
\begin{equation}\label{eq:kl_loss_expanded}
\begin{aligned}
    L_{\mathrm{KL}}^{(n)}
    ={}&\ln\!\left(
    \frac{\Gamma(\sum_{c=1}^{K}\tilde\alpha_c^{(n)})}
    {\Gamma(K)\prod_{c=1}^{K}\Gamma(\tilde\alpha_c^{(n)})}\right)\\
    &+\sum_{c=1}^{K}(\tilde\alpha_c^{(n)}-1)\\
    &\quad\times\left[
    \psi(\tilde\alpha_c^{(n)})
    -\psi(\sum_{c=1}^{K}\tilde\alpha_c^{(n)})\right].
\end{aligned}
\end{equation}

\subsection{Validity of Evidential State}
Softplus is strictly positive for finite real valued logits. Hence $\alpha_c=e_c+1>1$ and $S=\sum_c\alpha_c>K$. The expected probabilities obey $\hat{p}_c=\alpha_c/S>0$ and $\sum_c\hat{p}_c=1$. Thus $\hat{\bm{p}}\in\Delta^K$. Since $S>K$, $u_e=K/S\in(0,1)$ for finite logits. Its limiting value $1$ is approached when all evidence values $e_c$ tend to zero. Finally, $S\ge\alpha_c$ for every class and $\psi(\cdot)$ is monotone on positive inputs. Therefore $\psi(S+1)-\psi(\alpha_c+1)\ge0$. Each summand in Eq.~\ref{eq:u_a} is non negative. Hence $u_a\ge0$.

\subsection{2:4 Feasibility Preservation}
\label{app:24_feasibility}
For each eligible row $i$ and complete valid block $k$, Eq.~\ref{eq:mask} selects the two local indices in $I_{i,k}^{(l)}$ and assigns mask value $1$ only to those entries. The remaining two entries receive mask value $0$. Auxiliary padding entries may be used for bookkeeping. They have score $-\infty$, remain non trainable, and cannot consume active slots. Hence every complete valid 4-block contains exactly two active trainable entries. Coordinates outside complete valid blocks are excluded from the exact 2:4 claim and are reported through the valid block fraction. They are outside the numerator and denominator of the structured sparsity guarantee. The score projection objective decomposes over independent 4-blocks. Selecting the two largest scores in each block solves the projection exactly.

\subsection{Masked Gradient Isolation}
For each entry, $W_{\text{eff},ij}=W_{ij}M_{ij}$ with the mask fixed during the mini batch. By the chain rule,
\begin{equation}
\begin{aligned}
    \frac{\partial L}{\partial W_{ij}}
    &=\frac{\partial L}{\partial W_{\mathrm{eff},ij}}
    \frac{\partial W_{\mathrm{eff},ij}}{\partial W_{ij}}\\
    &=\frac{\partial L}{\partial W_{\mathrm{eff},ij}}M_{ij}.
\end{aligned}
\end{equation}
Stacking this entry wise identity over the tensor gives Eq.~\ref{eq:masked_gradient}. If $M_{ij}=0$, the instantaneous task gradient with respect to $W_{ij}$ is zero, which motivates the separate structural proxy pass used by the pruning/regrowth update.

\subsection{Virtual Effective Weight Probe for Dormant Saliency}
\label{app:virtual_probe_note}
Regrowth uses a dormant link saliency distinct from the ordinary parameter gradient $\partial L/\partial W_{ij}$ under a fixed mask. The structural refresh introduces an auxiliary dense effective tensor $\widetilde{\bm{W}}_{\text{eff}}$ initialized at $\bm{W}\odot\bm{M}$ and computes proxy gradients with respect to that tensor. For a dormant coordinate, $\widetilde{W}_{\text{eff},ij}=0$, yet $\partial L_{\text{grow}}/\partial \widetilde{W}_{\text{eff},ij}$ can still be nonzero. The coordinate is treated as a virtual effective input to the proxy computation. These gradients are discarded after ranking and are never applied as ordinary updates to dormant parameters. Masked gradient isolation for training and virtual probe saliency for topology scoring are compatible.

\subsection{Signed First Order Pruning Criterion}
\label{app:signed_pruning}
Let the mask be fixed during the structural proxy pass and consider an active effective coordinate $w_{\text{eff},ij}$. Define the removal perturbation $\delta_{ij}=-w_{\text{eff},ij}$. Taylor's theorem gives
\begin{equation}
\begin{aligned}
    R_{\mathcal B}(w_{\mathrm{eff},ij}+\delta_{ij})
    -R_{\mathcal B}(w_{\mathrm{eff},ij})
    &=\delta_{ij}\frac{\partial R_{\mathcal B}}
    {\partial w_{\mathrm{eff},ij}}\\
    &\quad+\mathcal R_{ij}.
\end{aligned}
\end{equation}
If the one dimensional second derivative along this coordinate is bounded by $L_{ij}$ on the segment between $w_{\text{eff},ij}$ and $0$, then
\[
    |\mathcal{R}_{ij}|
    \le
    \frac{L_{ij}}{2}|w_{\text{eff},ij}|^2.
\]
This gives
\begin{equation*}
\begin{aligned}
    R_{\mathcal B}(0)-R_{\mathcal B}(w_{\mathrm{eff},ij})
    &=-w_{\mathrm{eff},ij}\frac{\partial R_{\mathcal B}}
    {\partial w_{\mathrm{eff},ij}}\\
    &\quad+\mathcal R_{ij}.
\end{aligned}
\end{equation*}
For active links, $w_{\text{eff},ij}=w_{ij}$. The leading term is negative when
\begin{equation}\label{eq:proof_signed_positive}
    w_{\mathrm{eff},ij}
    \frac{\partial R_{\mathcal B}}{\partial w_{\mathrm{eff},ij}}>0.
\end{equation}
If the leading term dominates the curvature remainder, removal decreases the proxy risk ratio. Conversely, when
\begin{equation}\label{eq:proof_signed_negative}
    w_{\mathrm{eff},ij}
    \frac{\partial R_{\mathcal B}}{\partial w_{\mathrm{eff},ij}}<0.
\end{equation}
the leading term is positive and removal increases the risk ratio to first order. Otherwise, the sign remains a local saliency ranking rather than a deterministic descent certificate. The positive part operator in Eq.~\ref{eq:C_ij} keeps only the first case as a pruning candidate.

\subsection{Local Regrowth Saliency}
\label{app:regrowth_saliency}
Let $Q(\widetilde{\bm{W}}_{\text{eff}})=L_{\text{grow}}(\widetilde{\bm{W}}_{\text{eff}})$ denote the structural probe objective evaluated at the current masked effective tensor. For a scalar virtual perturbation $\delta_{ij}$ at connection $(i,j)$,
\begin{equation}
\begin{aligned}
    Q(\widetilde w_{\mathrm{eff},ij}+\delta_{ij})
    -Q(\widetilde w_{\mathrm{eff},ij})
    &=\delta_{ij}\frac{\partial Q}
    {\partial\widetilde w_{\mathrm{eff},ij}}\\
    &\quad+\mathcal R^Q_{ij}.
\end{aligned}
\end{equation}
If the coordinate wise curvature is bounded by $H_{ij}$ for $|\delta_{ij}|\le\eta$, then
\[
    |\mathcal{R}^{Q}_{ij}|
    \le
    \frac{H_{ij}}{2}\eta^2.
\]
Thus, to leading order, the largest achievable absolute perturbation under $|\delta_{ij}|\le\eta$ is
\[
    \eta\left|
    \frac{\partial Q}{\partial \widetilde{w}_{\text{eff},ij}}
    \right|.
\]
Ranking dormant candidates by Eq.~\ref{eq:G_ij} selects links with high local leading order influence on the KL uniform structural saliency objective. The sign is discarded. The score serves only candidate topology exploration before the valid block 2:4 projection.

\subsection{Bounded Morphological State}
The final operation in Eq.~\ref{eq:score_update} applies an element wise clamp to the interval $[-V_{\max},V_{\max}]$. Hence every updated score lies in that interval regardless of the magnitude of the unclamped momentum step. The mask is then recomputed from the bounded score tensor using Eq.~\ref{eq:mask}. Appendix~\ref{app:24_feasibility} applies at the next refresh.

\subsection{Controlled Imbalance Modulation}
Because $\operatorname{sg}[\hat{p}_{y_n}^{(n)}]$ is treated as constant during backpropagation, $\nabla_{\bm{e}}\mathcal{F}_n(t)=0$. Thus the sample level gradient is proportional to
\begin{equation}
\begin{aligned}
    \mathcal F_n(t)\nabla_{\bm e}\!\bigl(
    \bar\omega_{y_n}L_{\mathrm{CE}}^{(n)}
    &+\lambda_{\mathrm{KL}}\mu(t)\\
    &\quad\times L_{\mathrm{KL}}^{(n)}\bigr).
\end{aligned}
\end{equation}
Moreover, $\mathcal{F}_n(t)\ge0$, $\mu(t)\in[0,1]$, and $\bar{\omega}_{y_n}\le\omega_{\max}$ by Eq.~\ref{eq:sample_weight}. In the reported symmetric KL setting, the KL term has no additional class dependent multiplier. In the asymmetric diagnostic variant, the multiplier is capped at $\Lambda_{\max}$ by:
\begin{equation}\label{eq:asym_multiplier}
\begin{aligned}
    \Lambda_{\mathrm{asym}}^{(n)}
    &=\min(\bar\omega_{y_n},\Lambda_{\max}),\\
    \Lambda_{\max}&=10.0.
\end{aligned}
\end{equation}
Hence focal modulation changes sample emphasis without adding an uncontrolled confidence derivative path. The supervised class weight is explicitly capped.

\section{Novelty Boundary Against Closest Prior Art}
\label{app:novelty}
This appendix positions DST-EDL against adjacent method families. The purpose
is to make the novelty boundary explicit without asserting general superiority
over alternatives.

\begin{table}[tbp]
\centering
\caption{Compact novelty boundary against closest related methods. The distinction is Dirichlet derived proxy saliency used directly for dynamic valid block 2:4 topology adaptation with dormant link structural scoring.}
\label{tab:novelty}
\scriptsize
\renewcommand{\arraystretch}{1.05}
\setlength{\tabcolsep}{2pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}p{0.28\columnwidth}p{0.64\columnwidth}@{}}
\toprule
\textbf{Method family} & \textbf{Boundary vs. DST-EDL} \\
\midrule
SNIP~\cite{lee2019snip} & One shot connection sensitivity at initialization. No uncertainty guided dynamic topology. \\
SET~\cite{mocanu2018scalable} / SNFS~\cite{dettmers2019sparse} & Dynamic sparse training with random or task gradient driven regrowth. No Dirichlet ambiguity/vacuity driven regrowth. \\
RigL~\cite{evci2020rigging} / SRigL~\cite{lasby2023srigl} & Strong dynamic sparse baselines. No evidential ambiguity/vacuity controller for exact valid block 2:4 adaptation. \\
Fisher EDL~\cite{deng2023fisher}, R-EDL/Re-EDL~\cite{chen2023redl,chen2024reedl}, Flexible EDL~\cite{yoon2025flexible} & Refine evidential uncertainty modeling. No pruning, regrowth, or structured sparse feasibility. \\
NVIDIA 2:4 sparsity~\cite{mishra2021accelerating} & Provides a hardware compatible sparsity constraint, not uncertainty aware topology learning. \\
\textbf{DST-EDL (ours)} & \textbf{Uses Dirichlet ambiguity/vacuity and KL uniform proxy gradients for pruning/regrowth under exact valid block 2:4 feasibility, including dormant link structural scoring.} \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\section{Additional Empirical Scope}
\label{app:external}
The completed external-domain, softmax, and CIFAR-100-LT analyses below define
the empirical scope beyond the central ISIC comparison. Broader validation on
additional imbalance profiles, backbones, sparse Tensor Core compatible
kernels, and quality-gated failure detection remains outside the central claim.

\subsection{Held-out PAD-UFES-20 OOD Diagnostic}
PAD-UFES-20 contains smartphone clinical skin-lesion images and associated
metadata~\cite{pacheco2020padufes}. We evaluate source-only DST-EDL models on
the held-out \texttt{imgs\_part\_3} partition: no PAD image was used for OOD
exposure, model selection, calibration, or threshold selection. This diagnostic
uses the hardware-compatible deployment configuration and therefore does not
replace the predictive-evaluation ISIC results. With the layer-3 KNN score fixed as the
primary OOD score, the three models (seeds 42, 123, and 456) separate held-out
ISIC test images from the 728 available PAD images with AUROC
$0.9990\pm0.0001$ and AUPR $0.9380\pm0.0083$. All evaluated checkpoints
preserved the valid 2:4 sparse inference state.

This is a domain-separation result, not external diagnostic classification.
The accessible \texttt{imgs\_part\_3} images have one observed binary class,
so external classification AUROC/AP and Equalized Odds are undefined. We thus
do not make a PAD clinical-classification, subgroup-fairness, or OOD-exposure
improvement claim from this partition.

\subsection{ISIC Softmax Long Tailed Diagnostic}
Table~\ref{tab:isic_softmax_lt_diagnostic} reports the completed three seed
softmax long tailed baselines on the same ISIC split. These runs provide
context for ordinary probability based classifiers. The main comparison
centers on evidential models. Softmax baselines lack Dirichlet vacuity, expected
entropy, and valid block sparse topology signals. The strongest softmax row in
this diagnostic is MiSLAS, with pAUC$_{0.80}$ $0.1032\pm0.0064$ and
Macro AUROC $0.8708\pm0.0090$. These values remain below the reported DST-EDL means of
$0.1172$ and $0.8853$ under the evidential sparse protocol.

The diagnostic also shows a thresholding pathology typical of extreme prevalence
screening. Standard CE, Focal Loss, LDAM-DRW, cRT, and MiSLAS produce no
positive predictions at the default threshold across all three seeds
(default positive rate $0.0000$). Their default threshold balanced accuracy
collapses to $0.5000$ despite nontrivial ranking metrics. Class Balanced CE and
Balanced Softmax predict a small positive fraction at the default threshold.
They remain substantially lower in pAUC$_{0.80}$ and Macro AUROC. These results
motivate the paper's use of ranking metrics and validation selected operating
points instead of default threshold accuracy under the ISIC 2024 prevalence. The
very low softmax ECE values should also be interpreted with this prevalence
effect in mind. They primarily measure probability calibration under a negative
dominated test distribution rather than evidential uncertainty quality.

\begin{table}[tbp]
\centering
\scriptsize
\caption{ISIC 2024 softmax long tailed diagnostic over seeds 42, 123, and 456. Values are mean $\pm$ std. Default+ is the default threshold positive prediction rate.}
\label{tab:isic_softmax_lt_diagnostic}
\setlength{\tabcolsep}{1.5pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l C C C C C@{}}
\toprule
Model & pAUC$_{0.80}$ & Macro AUROC & AP & ECE & Default+ \\
\midrule
Standard CE & $0.0907\pm0.0078$ & $0.8560\pm0.0106$ & $0.0274\pm0.0018$ & $0.0007\pm0.0001$ & $0.0000\pm0.0000$ \\
Focal Loss & $0.0941\pm0.0137$ & $0.8585\pm0.0213$ & $0.0363\pm0.0082$ & $0.0015\pm0.0002$ & $0.0000\pm0.0000$ \\
Class Balanced CE & $0.0729\pm0.0107$ & $0.8314\pm0.0147$ & $0.0301\pm0.0036$ & $0.0028\pm0.0003$ & $0.0047\pm0.0002$ \\
Balanced Softmax & $0.0743\pm0.0119$ & $0.8366\pm0.0101$ & $0.0314\pm0.0034$ & $0.0017\pm0.0002$ & $0.0031\pm0.0003$ \\
LDAM-DRW & $0.0985\pm0.0063$ & $0.8623\pm0.0046$ & $0.0319\pm0.0052$ & $0.0009\pm0.0001$ & $0.0000\pm0.0000$ \\
cRT & $0.0905\pm0.0078$ & $0.8499\pm0.0137$ & $\mathbf{0.0424\pm0.0148}$ & $0.0009\pm0.0001$ & $0.0000\pm0.0000$ \\
MiSLAS & $\mathbf{0.1032\pm0.0064}$ & $\mathbf{0.8708\pm0.0090}$ & $0.0313\pm0.0051$ & $0.0015\pm0.0001$ & $0.0000\pm0.0000$ \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

\subsection{CIFAR-100-LT Diagnostic and Protocol}
Beyond the completed 1:100 diagnostic reported here, broader long tailed
evaluation should vary imbalance profiles (1:10, 1:50, 1:100) and report
top 1/top 5 accuracy, balanced accuracy, macro F1, many/medium/few shot accuracy,
macro AUROC, macro AP, NLL, Brier score, ECE, AURC, and failure detection
metrics against standard long tailed methods (CE, Focal Loss, Logit Adjustment,
Class Balanced Loss, Balanced Softmax, LDAM-DRW, cRT, and MiSLAS style LAS+cRT),
evidential baselines (Dense EDL, Fisher EDL, Flexible EDL, and R-EDL), and
sparse baselines (Static 2:4 EDL and RigL-style dynamic sparsity).

A completed CIFAR-100-LT 1:100 run over seeds 42, 123, and 456 is included as a diagnostic stress test. DST-EDL preserves 50\% active density and valid 2:4 blocks and obtains lower ECE than the dense CIFAR baselines. Standard CE and Dense EDL remain stronger on classification and ranking metrics. This negative result bounds the paper's scope: the present evidence supports an uncertainty-guided valid-block sparse mechanism in the ISIC screening setting, not a universal long-tailed classification recipe.

\begin{table}[tbp]
\centering
\scriptsize
\caption{CIFAR-100-LT 1:100 diagnostic over seeds 42, 123, and 456. Values are mean $\pm$ std. The run is reported as a scope and stress test result outside the central performance claim.}
\label{tab:cifar_ir100_diagnostic}
\setlength{\tabcolsep}{1.8pt}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}l C C C C@{}}
\toprule
Model & Bal. & Macro AUROC & ECE & AURC \\
\midrule
Standard CE & $0.4097\pm0.0061$ & $\mathbf{0.9409\pm0.0029}$ & $0.2578\pm0.0116$ & $\mathbf{0.3513\pm0.0093}$ \\
Dense EDL & $0.3623\pm0.0148$ & $0.7600\pm0.0080$ & $0.2321\pm0.0216$ & $0.4809\pm0.0409$ \\
DST-EDL & $0.3411\pm0.0080$ & $0.7524\pm0.0059$ & $\mathbf{0.2036\pm0.0074}$ & $0.5151\pm0.0170$ \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

For DST-EDL, all three CIFAR-100-LT runs preserved 50.0\% active density. They satisfied all valid 2:4 blocks in the wrapped layers (2,789,376/2,789,376). No representational collapse was detected. The long tail breakdown remains few shot limited. Mean many shot, medium shot, and few shot accuracies are 0.6004, 0.3063, and 0.0864, respectively. The diagnostic separates structural feasibility from task level competitiveness. Exact 2:4 validity remains intact. CIFAR-100-LT performance indicates that additional tuning or dataset specific design is needed before making broader long tailed claims.

\subsection{Hardware Evidence and Sparse-Backend Evaluation}
Hardware evidence separates structural feasibility, framework-level execution cost, and realized sparse-kernel acceleration. Prior 2:4 work reports that Sparse Tensor Cores can increase matrix-math throughput under strict 50\% structured sparsity~\cite{mishra2021accelerating}. Deployment libraries such as cuSPARSELt target matrix multiplication with a 50\% structured sparse operand~\cite{nvidiacusparselt}. PyTorch semi-structured tensors similarly require CUDA tensors with supported shape and dtype constraints~\cite{pytorchsemistructured}. Valid masks therefore establish compatibility, whereas speedup requires a compatible sparse-kernel path.

\begin{table}[tbp]
\centering
\scriptsize
\caption{Hardware evidence taxonomy. Structural validity, framework-level execution, and backend acceleration answer distinct empirical questions; realized acceleration requires the backend category.}
\label{tab:hardware_protocol_ladder}
\setlength{\tabcolsep}{1.4pt}
\renewcommand{\arraystretch}{1.08}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{@{}p{0.16\columnwidth}p{0.28\columnwidth}p{0.31\columnwidth}p{0.20\columnwidth}@{}}
\toprule
\textbf{Evidence class} & \textbf{Comparison} & \textbf{Measurements} & \textbf{Supported claim} \\
\midrule
Structural validity & Dense EDL, Static 2:4, RigL-style 2:4, DST-EDL & Active density, valid 2:4 blocks, wrapped layer parameter reduction, theoretical MAC and FLOP reduction & Sparse compatibility \\
Framework execution & Dense EDL, Static 2:4, DST-EDL & Images per second, milliseconds per batch, peak CUDA memory, batch size, image size, GPU name & Framework-level cost \\
Backend acceleration & Dense exported model vs. exported 2:4 model & cuSPARSELt, TensorRT, or PyTorch semi structured latency; throughput; memory; layout conversion status & Realized speedup when sparse kernels are selected \\
\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}

Structural validity and framework-level execution are assessed with \path{experiments/hardware_profile.py}. The analysis records active density, valid 2:4 block fraction, dense execution MACs, theoretical active MACs, and FLOPs under the two-MACs-per-multiply--add convention, together with masked-PyTorch throughput, latency, and peak CUDA memory. These measurements characterize feasibility and framework-level cost because masked kernels still multiply dense tensors without exposing a compressed 2:4 layout to Sparse Tensor Cores. Backend acceleration is evaluated separately with \path{experiments/nvidia_sparse_benchmark.py}, exported ONNX graphs, FP16 execution, \texttt{--sparsity=enable}, and build-log verification of sparse-tactic selection. This backend evidence supports the reported RTX A2000 batch-throughput speedup; operator, batch-size, and device dependence remain explicit scope conditions.

\subsection{Selective-Prediction Scope}
A quality-control filter based on the entropy-derived ambiguity proxy $u_a$ remains a future evaluation. Correlation with artifacts such as glare, motion blur, and obscuration, together with accepted-versus-deferred risk at fixed coverage, would test whether the proxy has selective-prediction utility.

CIFAR-100-LT separates general imbalance behavior from the ISIC screening setting. Hardware evidence likewise distinguishes structural 2:4 feasibility from realized acceleration, which requires compatible sparse Tensor Core kernels and layouts such as cuSPARSELt or TensorRT sparse execution. Quality-gated failure detection based on $u_a$ or related ambiguity scores remains future work and belongs to selective prediction rather than a clinical deployment claim.

\fi
\end{document}
