\bibitem[\protect\citeauthoryear{Guo \bgroup et al.\egroup }{2017}]{guo2017calibration}
Guo, C.; Pleiss, G.; Sun, Y.; and Weinberger, K. Q.. 2017. 
\newblock On calibration of modern neural networks. In {\em International Conference on Machine Learning (ICML)}.

\bibitem[\protect\citeauthoryear{Sensoy \bgroup et al.\egroup }{2018}]{sensoy2018evidential}
Sensoy, M.; Kaplan, L.; and Kandemir, M.. 2018. 
\newblock Evidential deep learning to quantify classification uncertainty. In {\em Advances in Neural Information Processing Systems (NeurIPS)}.

\bibitem[\protect\citeauthoryear{Malinin and Gales}{2018}]{malinin2018predictive}
Malinin, A.; and Gales, M.. 2018. 
\newblock Predictive uncertainty estimation via prior networks. In {\em Advances in Neural Information Processing Systems (NeurIPS)}.

\bibitem[\protect\citeauthoryear{Evci \bgroup et al.\egroup }{2020}]{evci2020rigging}
Evci, U.; Gale, T.; Menick, J.; Castro, P. S.; and Elsen, E.. 2020. 
\newblock Rigging the lottery: Making all tickets winners. In {\em International Conference on Machine Learning (ICML)}.

\bibitem[\protect\citeauthoryear{Cui \bgroup et al.\egroup }{2019}]{cui2019class}
Cui, Y.; Jia, M.; Lin, T.-Y.; Song, Y.; and Belongie, S.. 2019. 
\newblock Class balanced loss based on effective number of samples. In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}.

\bibitem[\protect\citeauthoryear{Cao \bgroup et al.\egroup }{2019}]{cao2019learning}
Cao, K.; Wei, C.; Gaidon, A.; Arechiga, N.; and Ma, T.. 2019. 
\newblock Learning imbalanced datasets with label distribution aware margin loss. In {\em Advances in Neural Information Processing Systems (NeurIPS)}.

\bibitem[\protect\citeauthoryear{Menon \bgroup et al.\egroup }{2021}]{menon2021long}
Menon, A. K.; Jayasumana, S.; Rawat, A. S.; Jain, H.; Veit, A.; and Kumar, S.. 2021. 
\newblock Long tail learning via logit adjustment. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Ren \bgroup et al.\egroup }{2020}]{ren2020balanced}
Ren, J.; Yu, C.; Sheng, S.; Ma, X.; Zhao, H.; Yi, S.; and Li, H.. 2020. 
\newblock Balanced meta softmax for long tailed visual recognition. In {\em Advances in Neural Information Processing Systems (NeurIPS)}.

\bibitem[\protect\citeauthoryear{Kang \bgroup et al.\egroup }{2020}]{kang2019decoupling}
Kang, B.; Xie, S.; Rohrbach, M.; Yan, Z.; Gordo, A.; Feng, J.; and Kalantidis, Y.. 2020. 
\newblock Decoupling representation and classifier for long tailed recognition. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Zhong \bgroup et al.\egroup }{2021}]{zhong2021improving}
Zhong, Z.; Cui, J.; Liu, S.; and Jia, J.. 2021. 
\newblock Improving calibration for long tailed recognition. In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}.

\bibitem[\protect\citeauthoryear{Zhou \bgroup et al.\egroup }{2020}]{zhou2020bbn}
Zhou, B.; Cui, Q.; Wei, X.-S.; and Chen, Z.-M.. 2020. 
\newblock BBN: Bilateral branch network with cumulative learning for long tailed visual recognition. In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}.

\bibitem[\protect\citeauthoryear{Cui \bgroup et al.\egroup }{2021}]{cui2021parametric}
Cui, J.; Zhong, Z.; Liu, S.; Yu, B.; and Jia, J.. 2021. 
\newblock Parametric contrastive learning. In {\em IEEE/CVF International Conference on Computer Vision (ICCV)}.

\bibitem[\protect\citeauthoryear{Deng \bgroup et al.\egroup }{2023}]{deng2023fisher}
Deng, D.; Chen, G.; Yu, Y.; Liu, F.; and Heng, P.-A.. 2023. 
\newblock Uncertainty estimation by Fisher information based evidential deep learning. In {\em International Conference on Machine Learning (ICML)}, PMLR.

\bibitem[\protect\citeauthoryear{Chen \bgroup et al.\egroup }{2024}]{chen2023redl}
Chen, M.; Gao, J.; and Xu, C.. 2024. 
\newblock R-EDL: Relaxing nonessential settings of evidential deep learning. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Chen \bgroup et al.\egroup }{2024}]{chen2024reedl}
Chen, M.; Gao, J.; and Xu, C.. 2024. 
\newblock Revisiting essential and nonessential settings of evidential deep learning. {\em arXiv preprint arXiv:2410.00393}.

\bibitem[\protect\citeauthoryear{Yoon and Kim}{2025}]{yoon2025flexible}
Yoon, T.; and Kim, H.. 2025. 
\newblock Uncertainty estimation by flexible evidential deep learning. {\em arXiv preprint arXiv:2510.18322}.

\bibitem[\protect\citeauthoryear{Yu \bgroup et al.\egroup }{2024}]{yu2024anedl}
Yu, Y.; Deng, D.; Liu, F.; Jin, Y.; Dou, Q.; Chen, G.; and Heng, P.-A.. 2024. 
\newblock ANEDL: Adaptive negative evidential deep learning for open set semi supervised learning. In {\em AAAI Conference on Artificial Intelligence}.

\bibitem[\protect\citeauthoryear{Wu \bgroup et al.\egroup }{2024}]{wu2024evidence}
Wu, Y.; Shi, B.; Dong, B.; Zheng, Q.; and Wei, H.. 2024. 
\newblock The evidence contraction issue in deep evidential regression: Discussion and solution. In {\em AAAI Conference on Artificial Intelligence}.

\bibitem[\protect\citeauthoryear{Shen \bgroup et al.\egroup }{2024}]{shen2024mirage}
Shen, M.; Ryu, J. J.; Ghosh, S.; Bu, Y.; Sattigeri, P.; Das, S.; and Wornell, G. W.. 2024. 
\newblock Are uncertainty quantification capabilities of evidential deep learning a mirage? In {\em Advances in Neural Information Processing Systems (NeurIPS)}.

\bibitem[\protect\citeauthoryear{Geifman and El-Yaniv}{2017}]{geifman2017selective}
Geifman, Y.; and El-Yaniv, R.. 2017. 
\newblock Selective classification for deep neural networks. In {\em Advances in Neural Information Processing Systems (NeurIPS)}.

\bibitem[\protect\citeauthoryear{Geifman and El-Yaniv}{2019}]{geifman2019selectivenet}
Geifman, Y.; and El-Yaniv, R.. 2019. 
\newblock SelectiveNet: A deep neural network with an integrated reject option. In {\em International Conference on Machine Learning (ICML)}.

\bibitem[\protect\citeauthoryear{Ding \bgroup et al.\egroup }{2020}]{ding2020revisiting}
Ding, Y.; Liu, J.; Xiong, J.; and Shi, Y.. 2020. 
\newblock Revisiting the evaluation of uncertainty estimation and its application to explore model complexity uncertainty tradeoff. In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition Workshops (CVPRW)}.

\bibitem[\protect\citeauthoryear{Traub \bgroup et al.\egroup }{2024}]{traub2024overcoming}
Traub, J.; Bungert, T. J.; L{\"u}th, C. T.; Baumgartner, M.; Maier-Hein, K. H.; Maier-Hein, L.; and Jaeger, P. F.. 2024. 
\newblock Overcoming common flaws in the evaluation of selective classification systems. In {\em Advances in Neural Information Processing Systems (NeurIPS)}.

\bibitem[\protect\citeauthoryear{Dettmers and Zettlemoyer}{2019}]{dettmers2019sparse}
Dettmers, T.; and Zettlemoyer, L.. 2019. 
\newblock Sparse networks from scratch: Faster training without losing performance. In {\em ICLR Workshop}.

\bibitem[\protect\citeauthoryear{Lee \bgroup et al.\egroup }{2019}]{lee2019snip}
Lee, N.; Ajanthan, T.; and Torr, P. H. S.. 2019. 
\newblock SNIP: Single shot network pruning based on connection sensitivity. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Li \bgroup et al.\egroup }{2025}]{li2025pffdst}
Li, L.; Yang, H.; Guo, J.; Yu, H.; Qin, M.; and Zhang, T.. 2025. 
\newblock An efficient and accurate dynamic sparse training framework based on parameter freezing. In {\em AAAI Conference on Artificial Intelligence}.

\bibitem[\protect\citeauthoryear{Lasby \bgroup et al.\egroup }{2024}]{lasby2023srigl}
Lasby, M.; Golubeva, A.; Evci, U.; Nica, M.; and Ioannou, Y.. 2024. 
\newblock Dynamic sparse training with structured sparsity. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Mishra \bgroup et al.\egroup }{2021}]{mishra2021accelerating}
Mishra, Asit; Latorre, Jorge Albericio; Pool, Jeff; Stosic, Darko; Stosic, Dusan; Venkatesh, Ganesh; Yu, Chong; and Micikevicius, Paulius. 2021. 
\newblock Accelerating sparse deep neural networks. {\em arXiv preprint arXiv:2104.08378}.

\bibitem[\protect\citeauthoryear{Hu \bgroup et al.\egroup }{2024}]{hu2024accelerating}
Hu, Y.; Zhao, K.; Huang, W.; Chen, J.; and Zhu, J.. 2024. 
\newblock Accelerating transformer pre-training with 2:4 sparsity. {\em arXiv preprint arXiv:2404.01847}.

\bibitem[\protect\citeauthoryear{NVIDIA}{2026}]{nvidiacusparselt}
NVIDIA. 2026. 
\newblock cuSPARSELt: A high-performance CUDA library for sparse matrix-matrix multiplication. \url{https://docs.nvidia.com/cuda/cusparselt/}, accessed.

\bibitem[\protect\citeauthoryear{PyTorch}{2026}]{pytorchsemistructured}
PyTorch. 2026. 
\newblock Semi-structured sparsity and \texttt{to\_sparse\_semi\_structured}. \url{https://docs.pytorch.org/docs/2.12/sparse.html}, accessed.

\bibitem[\protect\citeauthoryear{Mocanu \bgroup et al.\egroup }{2018}]{mocanu2018scalable}
Mocanu, D. C.; Mocanu, E.; Stone, P.; Nguyen, P. H.; Gibescu, M.; and Liotta, A.. 2018. 
\newblock Scalable training of artificial neural networks with adaptive sparse connectivity inspired by network science. {\em Nature Communications}, 9(1):2383.

\bibitem[\protect\citeauthoryear{Han \bgroup et al.\egroup }{2016}]{han2016deep}
Han, S.; Mao, H.; and Dally, W. J.. 2016. 
\newblock Deep compression: Compressing deep neural networks with pruning, trained quantization and Huffman coding. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Frankle and Carbin}{2019}]{frankle2019lottery}
Frankle, J.; and Carbin, M.. 2019. 
\newblock The lottery ticket hypothesis: Finding sparse, trainable neural networks. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Liu \bgroup et al.\egroup }{2019}]{liu2019large}
Liu, Z.; Miao, Z.; Zhan, X.; Wang, J.; Gong, B.; and Yu, S. X.. 2019. 
\newblock Large scale long tailed recognition in an open world. In {\em IEEE/CVF Conference on Computer Vision and Pattern Recognition (CVPR)}.

\bibitem[\protect\citeauthoryear{Collaboration}{2024}]{isic2024slice3d}
Collaboration, International Skin Imaging. 2024. 
\newblock SLICE-3D 2024 Challenge Dataset. International Skin Imaging Collaboration, 2024. doi:10.34970/2024-slice-3d. \url{https://doi.org/10.34970/2024-slice-3d}.

\bibitem[\protect\citeauthoryear{Jaeger \bgroup et al.\egroup }{2023}]{jaeger2023call}
Jaeger, P. F.; L{\"u}th, C. T.; Klein, L.; and Bungert, T. J.. 2023. 
\newblock A call to reflect on evaluation practices for failure detection in image classification. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Zhou \bgroup et al.\egroup }{2023}]{zhou2023domain}
Zhou, M.; Jamzad, A.; Izard, J.; Menard, A.; Siemens, R.; and Mousavi, P.. 2023. 
\newblock Domain transfer through image to image translation for uncertainty aware prostate cancer classification. {\em arXiv preprint arXiv:2307.00479}.

\bibitem[\protect\citeauthoryear{Micikevicius \bgroup et al.\egroup }{2018}]{micikevicius2018mixed}
Micikevicius, P.; Narang, S.; Alben, J.; Diamos, G.; Elsen, E.; Garcia, D.; Ginsburg, B.; Houston, M.; Kuchaiev, O.; Venkatesh, G.; and Wu, H.. 2018. 
\newblock Mixed precision training. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Loshchilov and Hutter}{2019}]{loshchilov2019decoupled}
Loshchilov, I.; and Hutter, F.. 2019. 
\newblock Decoupled weight decay regularization. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Kingma and Ba}{2015}]{kingma2015adam}
Kingma, D. P.; and Ba, J.. 2015. 
\newblock Adam: A method for stochastic optimization. In {\em International Conference on Learning Representations (ICLR)}.

\bibitem[\protect\citeauthoryear{Charpentier \bgroup et al.\egroup }{2020}]{charpentier2020posterior}
Charpentier, B.; Z{\"u}gner, D.; and G{\"u}nnemann, S.. 2020. 
\newblock Posterior Network: Uncertainty estimation without OOD samples via density based pseudo counts. In {\em Advances in Neural Information Processing Systems (NeurIPS)}.
